### 2020-05-13 paper/article_2.tex
struttura base paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{Article Title} % Article title
\author{%
\textsc{John Smith}\thanks{A thank you or further information} \\[1ex] % Your name
\normalsize University of California \\ % Your institution
\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
%\and % Uncomment if 2 authors are required, duplicate these 4 lines if more
%\textsc{Jane Smith}\thanks{Corresponding author} \\[1ex] % Second author's name
%\normalsize University of Utah \\ % Second author's institution
%\normalsize \href{mailto:jane@smith.com}{jane@smith.com} % Second author's email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent The recent tensions on the measured value of the Hubble constant between CMB and astrophisical observations, has triggered the need of new methods for its determination. In view of this, an effort has been done by H0LiCoW to use the gravitational lensing of quasars as a probe for H0. 
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

\blindtext % Dummy text

%------------------------------------------------

\section{Methods}

Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas. 
\begin{itemize}
\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
\item Curabitur feugiat
\item turpis sed auctor facilisis
\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
\item Fusce fermentum, mi sit amet euismod rutrum
\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
\item Pellentesque bibendum pretium aliquet
\end{itemize}
\blindtext % Dummy text

Text requiring further explanation\footnote{Example footnote}.

%------------------------------------------------

\section{Results}

\begin{table}
\caption{Example table}
\centering
\begin{tabular}{llr}
\toprule
\multicolumn{2}{c}{Name} \\
\cmidrule(r){1-2}
First name & Last Name & Grade \\
\midrule
John & Doe & $7.5$ \\
Richard & Miles & $2$ \\
\bottomrule
\end{tabular}
\end{table}

\blindtext % Dummy text

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\blindtext % Dummy text

%------------------------------------------------

\section{Discussion}

\subsection{Subsection One}

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

\subsection{Subsection Two}

\blindtext % Dummy text

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2020-05-13 paper/article_2.tex
struttura base paper

@@
 In view of this, an effort has been done by H0LiCoW to use the gravitational lensing of quasars as a probe for H0.
+This type of measurement requires a long term monitoring of lensed quasars (of the order of years).
+Since big telescopes have to deal with many observational requests, it is difficult to have a constant monitoring over the years, therefore this task can be achieved more easily by small/medium size telescopes.
+However, the number of lensed quasars with multiple images that can be resolved by these telescopes drops drastically.
+Here we present a method to deal with \textbf{non resolved} lensed quasars.
 \end{abstract}

### 2020-06-18 paper/article_2.tex
struttura base paper
updates

@@
 Here we present a method to deal with \textbf{non resolved} lensed quasars.
+This method has also the advantage of being less dependent on the microlensing effect of the lens galaxy.
 \end{abstract}
@@
 \section{Introduction}
-
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
-\blindtext % Dummy text
+In the last years, the precision of the Planck [cita] experiment, whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model $\Lambda$CDM which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
+In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$.
 

### 2020-06-18 paper/article_2.tex
updates

@@
 In the last years, the precision of the Planck [cita] experiment, whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model $\Lambda$CDM which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
-In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$.
+In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$, whose inference strongly depends on the assumed cosmological model.
+For example, relaxing the spatial flatness hypothesis of our Universe or the constant equation of state for the dark energy, would change the estimated value of $H_0$.
+\\
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars and, in the future, gravitational waves.
 

### 2020-06-18 paper/article_2.tex
updates

@@
 \\
-In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars and, in the future, gravitational waves.
-
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars [cita H0LiCOW] and, in the future, gravitational waves [cita].
+\\
+In this paper we will focus on the gravitational lensing method, which allows to measure $H_0$, as firstly suggested by Refsdal [cita], with a weaker dependence on the cosmological parameters if compared to the CMB analyses.
 %------------------------------------------------

### 2020-06-19 paper/article_2.tex
updates

@@
 \\
-In this paper we will focus on the gravitational lensing method, which allows to measure $H_0$, as firstly suggested by Refsdal [cita], with a weaker dependence on the cosmological parameters if compared to the CMB analyses.
+In this paper we will focus on the gravitational lensing method, which allows to measure $H_0$, as firstly suggested by Refsdal \cite{refsdal}, which has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
 %------------------------------------------------
@@
 
+\bibitem{refsdal}
+S. Refsdal,
+\textit{MNRAS, 128, 307} (1964)
+
 \bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}

### 2020-06-19 paper/article_2.tex
updates

--- 2020-06-18 paper/article_2.tex
@@
 \\
-In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars and, in the future, gravitational waves.
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars [cita H0LiCOW] and, gravitational waves [cita].
+\\
+In this paper we will focus on the gravitational lensing: firstly suggested by Refsdal \cite{refsdal}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
+This method depends on the matter distribution of both the lensing object (such as a galaxy) and along the line of sight, and it has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
+In particular, it depends on the matter density $\Omega_m$, the dark energy density $\Omega_\Lambda$, the curvature parameter $\Omega_k$ and the dark energy equation of state $\omega$ [cita].
+\\
+This method requires a long (of the order of years) photometric monitoring of the multiple images of the source and a good temporal sampling, to be able to observe the photometric variations of the source.
+In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil}.
 
@@
 
+\bibitem{refsdal}
+S. Refsdal,
+\textit{MNRAS, 128, 307} (1964)
+
+\bibitem{cosmograil}
+The COSMOGRAIL Collaboration,
+\textit{arXiv:2002.05736v1 [astro-ph.CO]} (2020)
+
 \bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}

### 2020-06-19 paper/article_2.tex
updates

@@
 \section{Introduction}
-In the last years, the precision of the Planck [cita] experiment, whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model $\Lambda$CDM which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
+In the last years, the precision of the Planck experiment [cita], whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model $\Lambda$CDM which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
 In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$, whose inference strongly depends on the assumed cosmological model.
@@
 In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil}.
+And the H0LiCOW collaboration has used part of these data to evaluate $H_0$ with a precision of $2.4\%$ \cite{h0licow_XIII}.
 
@@
 
+\bibitem{h0licow_XIII}
+The H0LiCOW Collaboration,
+\textit{MNRAS, stz3094} (2020)
+
 \bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}

### 2020-06-19 paper/article_2.tex
updates

@@
 In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars [cita H0LiCOW] and, gravitational waves [cita].
+The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurement.
+This tension has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration, which agree with the distance ladder measured value of  $H_0$ and brings to a $5.7 \sigma$ tension with the CMB result.
 \\
@@
 This method requires a long (of the order of years) photometric monitoring of the multiple images of the source and a good temporal sampling, to be able to observe the photometric variations of the source.
-In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil}.
+In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil} with 1-2 m size telescopes.
 And the H0LiCOW collaboration has used part of these data to evaluate $H_0$ with a precision of $2.4\%$ \cite{h0licow_XIII}.
+\\
 

### 2020-06-19 paper/article_2.tex
updates

--- 2020-06-19 paper/article_2.tex
@@
 \section{Introduction}
-In the last years, the precision of the Planck experiment [cita], whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model $\Lambda$CDM which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
+In the last years, the precision of the Planck experiment [cita], whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model ($\Lambda$CDM) which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
 In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$, whose inference strongly depends on the assumed cosmological model.
-For example, relaxing the spatial flatness hypothesis of our Universe or the constant equation of state for the dark energy, would change the estimated value of $H_0$.
+For example, relaxing the spatial flatness hypothesis of our Universe or the constant equation of state for the dark energy, would impact the $H_0$ estimation.
 \\
-In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars [cita H0LiCOW] and, gravitational waves [cita].
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars \cite{h0licow_I} and, gravitational waves [cita].
+\\
+The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurements, which has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration \cite{h0licow_XIII}, whose measured value is $H_0=73.3^{+1.7}_{-1.8}$ km s$^{-1}$ Mpc$^{-1}$, in agreement with the distance ladder results and, together with them, with a $5.3 \, \sigma$ tension with the Planck analysis assuming flat $\Lambda$CDM.
 \\
 In this paper we will focus on the gravitational lensing: firstly suggested by Refsdal \cite{refsdal}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
-This method depends on the matter distribution of both the lensing object (such as a galaxy) and along the line of sight, and it has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
+This method depends on the matter distribution in the source light trajectory, namely the lensing object (such as a galaxy) and objects along the line of sight, and it has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
 In particular, it depends on the matter density $\Omega_m$, the dark energy density $\Omega_\Lambda$, the curvature parameter $\Omega_k$ and the dark energy equation of state $\omega$ [cita].
 \\
-This method requires a long (of the order of years) photometric monitoring of the multiple images of the source and a good temporal sampling, to be able to observe the photometric variations of the source.
-In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil}.
+This method requires a long photometric monitoring of the multiple images of the source, of the order of years, and a good temporal sampling, to be able to observe the photometric variations of the source.
+In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil} with 1-2 m size telescopes.
 And the H0LiCOW collaboration has used part of these data to evaluate $H_0$ with a precision of $2.4\%$ \cite{h0licow_XIII}.
+\\
+Improving the precision in the $H_0$ evaluation will help in finding the reason of this big discrepancy, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{weinberg_euclid}.
 
@@
 
+\bibitem{h0licow_I}
+The H0LiCOW Collaboration,
+\textit{MNRAS, Volume 468, Issue 3} (2017)
+
+\bibitem{h0licow_XIII}
+The H0LiCOW Collaboration,
+\textit{MNRAS, stz3094} (2020)
+
 \bibitem{refsdal}
@@
 The COSMOGRAIL Collaboration,
-\textit{arXiv:2002.05736v1 [astro-ph.CO]} (2020)
+\textit{arXiv:2002.05736v1} (2020)
 
-\bibitem{h0licow_XIII}
-The H0LiCOW Collaboration,
-\textit{MNRAS, stz3094} (2020)
-
-\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-\newblock Assortative pairing and life history strategy - a cross-cultural
-  study.
-\newblock {\em Human Nature}, 20:317--330.
+\bibitem{weinberg_euclid}
+D. Weinberg et al.,
+\textit{Phys. Rep., 530, 87} (2013)
 

### 2020-07-07 our_qso_code/regression.py
Added two modules

@@ -1,0 +1,35 @@
+import numpy as np
+from sklearn.gaussian_process import GaussianProcessRegressor
+
+
+def fit_GP_to_lightcurve(t, y, err_y, kernel):
+    gp = GaussianProcessRegressor(kernel=kernel, alpha=err_y ** 2,
+                                  n_restarts_optimizer=10, optimizer='fmin_l_bfgs_b', normalize_y=True)
+    gp.fit(np.expand_dims(t, 1), y)
+    return gp
+
+
+def numeric_derivative(f: np.ndarray, step: float) -> np.ndarray:
+    return (f[1:] - f[:-1]) / step
+
+
+def WAV(f, sigma, step):
+    f_prime = numeric_derivative(f, step)
+    weights = 2 / (sigma[:-1] + sigma[1:])
+    WAV = np.dot(np.abs(f_prime), weights) / weights.sum()
+    return WAV
+
+
+def time_delay_grid_search(y_pred_1: np.ndarray, y_pred_2: np.ndarray,
+                           sigma1: np.ndarray, sigma2: np.ndarray,
+                           gp_time_step: float,
+                           dt_min: float = 0, dt_max: float = 100) -> float:
+    shift = np.arange(int(dt_min/gp_time_step), int(dt_max/gp_time_step), 1)
+    win = len(shift)
+    WAV_values = []
+    for i in shift:
+        diff = y_pred_1[win:-win] - y_pred_2[win + i:-win + i]
+        sigma_diff = sigma1[win:-win] + sigma2[win + i:-win + i]
+        WAV_values.append(WAV(diff, sigma_diff, gp_time_step))
+    return np.argmin(WAV_values) * gp_time_step
+

### 2020-07-07 our_qso_code/structure_func.py

@@ -1,0 +1,27 @@
+from scipy import stats
+import numpy as np
+from typing import Tuple
+
+
+def compute_lags_matrix(t) -> np.ndarray:
+    N = len(t)
+    t_row_repeated = np.repeat(t[np.newaxis, :], N, axis=0)
+    t_col_repeated = np.repeat(t[:, np.newaxis], N, axis=1)
+    tau = np.abs(t_row_repeated - t_col_repeated)
+    return tau
+
+
+def estimate_structure_func_from_data(t, y, err_y) -> Tuple[np.ndarray]:
+    N = len(t)
+    y_row_repeated = np.repeat(y[np.newaxis, :], N, axis=0)
+    y_col_repeated = np.repeat(y[:, np.newaxis], N, axis=1)
+    err_y_row_repeated = np.repeat(err_y[np.newaxis, :], N, axis=0)
+    err_y_col_repeated = np.repeat(err_y[:, np.newaxis], N, axis=1)
+    v = (y_row_repeated - y_col_repeated)**2 - (err_y_row_repeated**2 + err_y_col_repeated**2)
+    tau = compute_lags_matrix(t)
+    tau_v_sorted = np.array([[mytau, myv] for mytau, myv in sorted(zip(tau.ravel(), v.ravel()))])
+    tau_vals = tau_v_sorted[:, 0]
+    v_vals = tau_v_sorted[:, 1]
+    tau_binned = stats.binned_statistic(tau_vals, tau_vals, bins=100)[0]
+    v_binned = stats.binned_statistic(tau_vals, v_vals, bins=100)[0]
+    return tau_binned, v_binned

### 2020-07-07 modules/data_generator.py
modifications

@@ -15,8 +15,8 @@ class DataGenerator:
     :type t_max: float
     :param delay: value of the time delay, unit is days
     :type delay: float
-    :param n_samples: number of samples
-    :type n_samples: int
+    :param sampling_rate: sampling rate in days^-1
+    :type sampling_rate: float
     :param seed: seed of numpy random generator
     :type seed: float
     :param n_gaps: number of gaps to simulate in artificial data
@@ -34,8 +34,8 @@ class DataGenerator:
     def __init__(self,
                  t_max=None,
                  delay=None,
-                 n_samples=None,
-                 image_ratio=1. / 1.44,
+                 sampling_rate=None,
+                 image_ratio=1.0,
                  seed=0,
                  n_gaps=None,
                  gap_size=None,
@@ -48,12 +48,7 @@ class DataGenerator:
         if delay is None:
             raise ParameterError("Time delay cannot be None")
         self.delay = delay
-        if n_samples is None or n_samples < 0 or n_samples == 0:
-            raise ParameterError("Number of samples cannot be None, negative or zero")
-        try:
-            self.n_samples = int(n_samples)
-        except:
-            raise Exception(f"Cannot cast variable {n_samples} as an integer")
+        self.sampling_rate = sampling_rate
         if relative_err is None or relative_err < 0:
             raise ParameterError("Relative error cannot be none or negative")
         self.relative_err = relative_err
@@ -62,6 +57,7 @@ class DataGenerator:
         self.image_ratio = image_ratio
 
         np.random.seed(seed)
+        self.n_samples = int(self.t_max * self.sampling_rate)
         self.t_regular = np.linspace(0, self.t_max, num=self.n_samples)
         self.t_domain = self.t_regular.copy()
         self.n_gaps = n_gaps
@@ -89,7 +85,7 @@ class DataGenerator:
         return means, sigmas, heights
 
     def generate_noisy_time_domain(self):
-        sampling_noise = (np.random.random(len(self.t_regular)) - 0.5) * (self.t_max / self.n_samples)
+        sampling_noise = (np.random.random(self.n_samples) - 0.5) * (self.t_max / self.n_samples)
         t_domain = self.t_regular + sampling_noise
         self.t_domain = t_domain.copy()
         return t_domain

### 2020-07-07 modules/example_dataset.py

@@ -11,7 +11,7 @@ gap_size = 3
 relative_err = 0.02
 
 datagen = dg.DataGenerator(t_max=t_max,
-                           n_samples=n_samples,
+                           sampling_rate=n_samples,
                            delay=delay,
                            seed=seed,
                            image_ratio=image_ratio,

### 2020-07-09 tex_files/artificial_data.tex
Implemented test main to do GP regression validation

\section{Structure function method}

\subsection{Structure function estimation from data}

Like in \cite{Press1992} one has to estimate from the data the so-called structure function

\begin{equation}
V(\tau) \equiv \frac{1}{2} \left \langle \left [ s(t+\tau) - s(t) \right ]^2  \right \rangle
\end{equation}

This is related to the covariance (or autocorrelation function) of the time serie (see \cite{Press1992}) for all details)

\begin{equation}
C(\tau) = \langle s(t)^2 \rangle - V(\tau)
\end{equation}

\subsection{Generating light curves from data}

\subsubsection{Eigenvalue decomposition}

Consider a given collection of $N$ observation times $t_1, \dots , t_N$. Construct the augmented observation time list appending to the end a the collection of the delayed observation times:

\[
t_1, \dots , t_N, t_1 - \Delta t, \dots, t_N - \Delta t
\]

with an assumed known delay $\Delta t$. After that use the estimated structure function to build a $2N \times 2N$ covariance matrix

\begin{equation}
C_{ij} = \langle s(t)^2 \rangle - V(\tau_{ij})
\end{equation}

with $\tau_{ij}$ being the lag matrix defined by

\begin{equation}
\tau_{ij} = |t_i - t_j|, \quad i,j = 1, \dots , 2N
\end{equation}

with the relation

\begin{equation}
t_{i+N} = t_i - \Delta t
\end{equation}

What they say in \cite{Press1992} is to perform the eigenvalue decomposition of the covariance matrix $C$

\begin{equation}\label{eq:eigen-cov-decomp}
C = Z^\dagger \Lambda Z
\end{equation}

with $\Lambda$ being the diagonal matrix with the eigenvalues

\[
\Lambda = \mathrm{diag} \left ( \lambda_1, \dots, \lambda_{2N} \right )
\]

 and $Z$ the orthogonal matrix which diagonalize $C$, the rows of $Z$ being the eigenvectors of $C$. After that one can generate a dataset with the following prescription

\begin{equation}
y_i = \sum_{i=1}^{2N} Z_{ij} \sqrt{\lambda_j} r_j + \langle n_i^2 \rangle^{1/2} r_i
\end{equation}

with $ \langle n_i^2 \rangle = e_i^2$ the squares of the errors on the data, and $r_i$, $r_j$ are independent uncorrelated gaussian random numbers with zero mean and unit variance. However we have tried this method on real data, and what it happens is that is not numerically stable. The eigenvalues span many orders of magnitude and this causes data with very high unrealistic peaks.

\subsubsection{Cholesky and Eigenvalue decompositions}

Assuming that $C$ is a real symmetric and positive definite matrix, it can be Cholesky decomposed

\begin{equation}
C = L L^{\dagger}
\end{equation}

with $L$ being a lower triangular matrix called the \emph{Cholesky factor} of $C$. We can show that there is a relation between the CHolseky and eigenvalue factorizations of a given matrix. First observe that, from the eigenvalue decomposition:

\begin{align*}
C = Z^\dagger \Lambda Z = Z^\dagger \sqrt{\Lambda} \sqrt{\Lambda} Z = \left ( \sqrt{\Lambda} Z \right )^{\dagger} \sqrt{\Lambda} Z
\end{align*}

with $\sqrt{\Lambda}$ being a diagonal matrix that multiplied by itself gives $\Lambda$, i.e. a diagonal matrix with the square roots of the eigenvalues as its entries. Now remember that any real square matrix can be factorized into a unitary matrix $Q$ and an upper triangular matrix $R$, the so-called $QR$-factorization. Doing this on the matrix $\sqrt{\Lambda} Z$ we have

\begin{equation}
\sqrt{\Lambda} Z = Q R
\end{equation}

But now we can rerite the \eqref{eq:eigen-cov-decomp} as

\[
C = (QR)^{\dagger} QR = R^{\dagger} Q^{\dagger} Q R = R^{\dagger} R 
\]

being $Q$ unitary. Finally, since $R$ is upper triangular, it conjugate transposed

\begin{align*}
L = R^{\dagger}
\end{align*}

is lower triangular. Therefore we have obtained the CHolesky decompotision of $C$

\begin{equation}
 C = L L^{\dagger}
\end{equation}

with

\[
L = \left ( \sqrt{\Lambda} Z \right )^{\dagger} Q
\]

\subsubsection{Reformulating the prescription}

Using the results of the previous section we can rewrite the prescription for the data generation (leave outside the noise for now)

\begin{equation}
\bv{y} = \left ( Z \sqrt{\Lambda} \right ) \bv{r} 
\end{equation}

Instead of using $Z$ use $Z^{\dagger}$ and obtain

\begin{align*}
Z^{\dagger} \sqrt{\Lambda} = \left ( \sqrt{\Lambda} Z\right )^{\dagger} = L Q^{\dagger}
\end{align*}

The we can substitute the eigenvalue decomposition with the Cholesky one, which is numerically more stable. For simplicity we simply write

\begin{equation}
\bv{y} = L \bv{r}
\end{equation}

### 2020-07-09 main_scripts/validate_gp_regression.py

@@ -1,0 +1,36 @@
+import h5py
+import numpy as np
+from modules import regression as rg
+import pandas as pd
+import time
+
+
+def main():
+    data = h5py.File('../aux/RXJ1131_random_delays_0_30.h5', 'r')
+    t_dom = data['t_domain'][()]
+    # A = data['original_data']['A'][()]
+    errA = data['original_data']['errA'][()]
+    realization_keys = [k for k in data.keys() if k.startswith('realization')]
+    true_delays = []
+    estimated_delays = []
+    for k in realization_keys:
+        y1 = data[k]['y'][()]
+        y2 = data[k]['y_delayed'][()]
+        true_delay = data[k]['delay'][()]
+
+        t0 = time.time()
+        estimated_delay = rg.estimate_delay(t_dom, y1, y2, errA, errA)
+        tf = time.time()
+        print(f'Elapsed time {tf-t0} s')
+
+        true_delays.append(true_delay)
+        estimated_delays.append(estimated_delay)
+
+    with open('results.txt', 'w') as f:
+        f.write('true_delay \t estimated_delay\n')
+        for td, ed in zip(true_delays, estimated_delays):
+            f.write(f'{td:.2f} \t {ed:.2f}\n')
+
+
+if __name__ == '__main__':
+    main()

### 2020-07-09 modules/regression.py

@@ -1,5 +1,6 @@
 import numpy as np
 from sklearn.gaussian_process import GaussianProcessRegressor
+from sklearn.gaussian_process.kernels import RBF, Matern, ConstantKernel
 
 
 def fit_GP_to_lightcurve(t, y, err_y, kernel):
@@ -32,4 +33,31 @@ def time_delay_grid_search(y_pred_1: np.ndarray, y_pred_2: np.ndarray,
         sigma_diff = sigma1[win:-win] + sigma2[win + i:-win + i]
         WAV_values.append(WAV(diff, sigma_diff, gp_time_step))
     return np.argmin(WAV_values) * gp_time_step
+
+
+def estimate_delay(t, y1, y2, err_y1, err_y2):
+    kernel = ConstantKernel(2, (1e-3, 1e2)) * Matern(length_scale=200.0, length_scale_bounds=(1, 300), nu=1.5)
+    gp1 = fit_GP_to_lightcurve(t, y1, err_y1, kernel)
+    gp2 = fit_GP_to_lightcurve(t, y2, err_y2, kernel)
+
+    gp_step_days = 0.2
+    support = np.arange(t[0] - 5e1, t[-1] + 5e1, gp_step_days)
+
+    y_pred_1, sigma1 = gp1.predict(np.expand_dims(support, 1), return_std=True)
+    y_pred_2, sigma2 = gp2.predict(np.expand_dims(support, 1), return_std=True)
+
+    dt_min_days = 0
+    dt_max_days = 50
+
+    shift = np.arange(int(dt_min_days / gp_step_days),
+                      int(dt_max_days / gp_step_days), 1)
+    win = len(shift)
+    WAV_values = []
+    for i in shift:
+        diff = y_pred_1[win:-win] - y_pred_2[win + i:-win + i]
+        sigma_diff = sigma1[win:-win] + sigma2[win + i:-win + i]
+        WAV_values.append(WAV(diff, sigma_diff, gp_step_days))
+    return np.argmin(WAV_values) * gp_step_days
+
+
 

### 2020-07-09 modules/structure_func.py

@@ -25,3 +25,5 @@ def estimate_structure_func_from_data(t, y, err_y) -> Tuple[np.ndarray]:
     tau_binned = stats.binned_statistic(tau_vals, tau_vals, bins=100)[0]
     v_binned = stats.binned_statistic(tau_vals, v_vals, bins=100)[0]
     return tau_binned, v_binned
+
+

### 2020-07-09 main_scripts/validate_gp_regression.py
changed outfile path

@@ -26,7 +26,7 @@ def main():
         true_delays.append(true_delay)
         estimated_delays.append(estimated_delay)
 
-    with open('results.txt', 'w') as f:
+    with open('../aux/results.txt', 'w') as f:
         f.write('true_delay \t estimated_delay\n')
         for td, ed in zip(true_delays, estimated_delays):
             f.write(f'{td:.2f} \t {ed:.2f}\n')

### 2020-07-09 main_scripts/validate_gp_regression.py
Workaround

@@ -3,6 +3,9 @@ import numpy as np
 from modules import regression as rg
 import pandas as pd
 import time
+import sys
+
+sys.path.insert(0, '..')
 
 
 def main():

### 2020-07-09 main_scripts/validate_gp_regression.py
Workaround v2

@@ -1,11 +1,12 @@
 import h5py
 import numpy as np
-from modules import regression as rg
 import pandas as pd
 import time
 import sys
 
 sys.path.insert(0, '..')
+
+from modules import regression as rg
 
 
 def main():

### 2020-07-09 main_scripts/validate_gp_regression.py
Adde prints

@@ -21,10 +21,12 @@ def main():
         y1 = data[k]['y'][()]
         y2 = data[k]['y_delayed'][()]
         true_delay = data[k]['delay'][()]
+        print(f'True delay: {true_delay}')
 
         t0 = time.time()
         estimated_delay = rg.estimate_delay(t_dom, y1, y2, errA, errA)
         tf = time.time()
+        print(f'Estimated delay: {estimated_delay}')
         print(f'Elapsed time {tf-t0} s')
 
         true_delays.append(true_delay)

### 2020-07-09 main_scripts/validate_gp_regression.py
Added logger

@@ -3,13 +3,19 @@ import numpy as np
 import pandas as pd
 import time
 import sys
+import logging
 
 sys.path.insert(0, '..')
 
 from modules import regression as rg
+from modules.utils import configure_logger
+
+logger = logging.getLogger()
 
 
 def main():
+    logfile = 'validate_GP_regression.log'
+    configure_logger(logger, logfile)
     data = h5py.File('../aux/RXJ1131_random_delays_0_30.h5', 'r')
     t_dom = data['t_domain'][()]
     # A = data['original_data']['A'][()]

### 2020-07-09 modules/utils.py

@@ -1,0 +1,23 @@
+import logging
+
+
+def configure_logger(logger, logfile):
+    logger.setLevel(logging.INFO)
+    # create file handler which logs even debug messages
+    fh = logging.FileHandler(logfile, mode='w')
+    fh.setLevel(logging.INFO)
+    # create stdout handler
+    stdout_handler = logging.StreamHandler(sys.stdout)
+    stderr_handler = logging.StreamHandler(sys.stderr)
+    stdout_handler.setLevel(logging.INFO)
+    stderr_handler.setLevel(logging.ERROR)
+    # create formatters and add them to the handlers
+    logfile_formatter = logging.Formatter('%(asctime)s - %(name)s - %(levelname)s - %(message)s')
+    fh.setFormatter(logfile_formatter)
+    stdout_handler.setFormatter(logging.Formatter('%(message)s'))
+    stderr_handler.setFormatter(logging.Formatter('%(message)s'))
+    # add the handlers to the logger
+    logger.addHandler(fh)
+    logger.addHandler(stdout_handler)
+    logger.addHandler(stderr_handler)
+

### 2020-07-09 modules/utils.py
bug

@@ -1,4 +1,5 @@
 import logging
+import sys
 
 
 def configure_logger(logger, logfile):

### 2020-07-09 main_scripts/validate_gp_regression.py
bug v2

@@ -27,13 +27,13 @@ def main():
         y1 = data[k]['y'][()]
         y2 = data[k]['y_delayed'][()]
         true_delay = data[k]['delay'][()]
-        print(f'True delay: {true_delay}')
+        logger.info(f'True delay: {true_delay}')
 
         t0 = time.time()
         estimated_delay = rg.estimate_delay(t_dom, y1, y2, errA, errA)
         tf = time.time()
-        print(f'Estimated delay: {estimated_delay}')
-        print(f'Elapsed time {tf-t0} s')
+        logger.info(f'Estimated delay: {estimated_delay}')
+        logger.info(f'Elapsed time {tf-t0} s')
 
         true_delays.append(true_delay)
         estimated_delays.append(estimated_delay)

### 2020-07-10 tex_files/artificial_data.tex
Added PRH class

@@
 with $L$ being a lower triangular matrix called the \emph{Cholesky factor} of $C$.
-We can show that there is a relation between the CHolseky and eigenvalue factorizations of a given matrix.
+We can show that there is a relation between the Cholesky and eigenvalue factorizations of a given matrix.
 First observe that, from the eigenvalue decomposition:
@@
 
-But now we can rerite the \eqref{eq:eigen-cov-decomp} as
+But now we can rewrite the \eqref{eq:eigen-cov-decomp} as
 
@@
 \end{equation}
+
+For what regards the noise, in ref \cite{Press1992} they identify the expectation value $\langle n_i^2 \rangle$ with the square of the quoted error on the magnitude.
+This makes sense if one assumes that $n_i$ is a random gaussian number with zero mean and variance given by the experimental error.
+So
+
+\begin{equation}
+\langle n_i^2 \rangle = e_i^2
+\end{equation}
+
+Then our prescription to generate the dataset becomes
+
+\begin{align}\label{eq:reform-artificial-prescr}
+y_i = \sum_{j=1}^{2N} L_{ij} r_j + e_i r_i, \quad i = 1, \dots , 2N
+\end{align}
+
+where $L$ is the lower triangular Cholesky factor of the estimated covariance $C$, i.e.
+
+\[
+C = L L^{\dagger}
+\]
+
+We interpret the first $N$ points $y_1, \dots, y_N$ as measurement of a mock light curve at times $t_1, \dots , t_N$, 
+while the other $N$ points $y_{N+1}, \dots, y_{2N}$ are measurements of the \emph{same} mock curve (apart from noise) at times 
+$t_1-\Delta t, \dots , t_N-\Delta t$, where the true delay can be decided \emph{a priori} when building the covariance.
+Therefore, if we assign each point of the second curve at times $t_1, \dots , t_N$ it will exactly experience a delay $\Delta t$.
+
+In figure \ref{fig:artif-dataset-5-vs-real} we see a mock light curve compared to the true light curve from which it has been generated.
+Both the curves have been scaled to have zero mean and unit variance, in order to make a direct comparison.
+In figure \ref{fig:artif-dataset-5-shift-comparison} we can see a comparison between the mock curves of the same mock dataset, with a relative time delay of about $18$ days.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.45]{realiz5_comparison.pdf}
+\caption{Realization 5 vs true quasar data, image A.}
+\label{fig:artif-dataset-5-vs-real}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.45]{Montecarlo_y_vs_y_delayed.pdf}
+\caption{Realization 5: mock light curve vs its shifted version, with a true delay of about $18$ days.}
+\label{fig:artif-dataset-5-shift-comparison}
+\end{figure}

### 2020-07-10 modules/prh.py

@@ -1,0 +1,178 @@
+import numpy as np
+import pandas as pd
+import itertools
+import h5py
+import matplotlib.pyplot as plt
+from scipy import stats
+from scipy import integrate
+from scipy import interpolate
+from scipy import signal
+from scipy.optimize import curve_fit
+import time
+from pathlib import Path
+import os
+from typing import Dict, Union, Callable, Tuple
+
+
+def power_law_sf(tau: Union[float, np.ndarray], **kwargs) -> Union[float, np.ndarray]:
+    return 10 ** kwargs['intercept'] * tau ** kwargs['slope']
+
+
+def exponential_sf(tau: Union[float, np.ndarray], **kwargs) -> Union[float, np.ndarray]:
+    return kwargs['V0'] * (1 - np.exp(- tau/kwargs['dtau0']))
+
+
+# noinspection PyTupleAssignmentBalance
+class LightCurvePRHGenerator:
+    t: np.ndarray
+    qso_lc_vals: np.ndarray
+    qso_lc_errs: np.ndarray
+    qso_df: pd.DataFrame
+    image: str
+    lags_matrix: np.ndarray
+    sf_lags: np.ndarray
+    sf_vals: np.ndarray
+    sf_fit_pars: Dict[str, float]
+    sf_fit_func: Callable
+
+    def __init__(self, image: str = 'A', sf_model: str = 'power_law', seed: float = 1234):
+        self.t = None
+        self.N_samples = None
+        self.qso_lc_vals = None
+        self.qso_lc_errs = None
+        self.qso_df = None
+        self.qso_id = None
+        self.image = image
+        self.sf_model = sf_model
+        self.sf_lags = None
+        self.sf_vals = None
+        self.sf_fit_pars = None
+        self.sf_fit_func = None
+        self.seed = seed
+        self.h5_dataset = None
+
+        if self.sf_model not in {'power_law', 'exponential'}:
+            raise ValueError(f'Invalid structure function model {self.sf_model}, available '
+                             f'options are: "power_law" or "exponential"')
+        np.random.seed(self.seed)
+
+    def loadQSODataFromFile(self, input_file: Union[str, Path]) -> None:
+        qso_df = pd.read_table(input_file)
+        self.t = qso_df['mhjd'].values
+        self.N_samples = len(self.t)
+        self.qso_df = qso_df
+        self.qso_id = Path(input_file).name.split('_')[0]
+        self.qso_lc_vals = self.qso_df[f'mag_{self.image}'].values
[…]

### 2020-07-11 modules/prh.py
Implemented dataset class and wrote example notebook

@@ -11,15 +11,15 @@ from scipy.optimize import curve_fit
 import time
 from pathlib import Path
 import os
-from typing import Dict, Union, Callable, Tuple
+from typing import Dict, Union, Callable, Tuple, List
 
 
-def power_law_sf(tau: Union[float, np.ndarray], **kwargs) -> Union[float, np.ndarray]:
-    return 10 ** kwargs['intercept'] * tau ** kwargs['slope']
+def power_law_sf(tau: Union[float, np.ndarray], pars) -> Union[float, np.ndarray]:
+    return 10 ** pars['intercept'] * tau ** pars['slope']
 
 
-def exponential_sf(tau: Union[float, np.ndarray], **kwargs) -> Union[float, np.ndarray]:
-    return kwargs['V0'] * (1 - np.exp(- tau/kwargs['dtau0']))
+def exponential_sf(tau: Union[float, np.ndarray], pars) -> Union[float, np.ndarray]:
+    return pars['V0'] * (1 - np.exp(- tau / pars['dtau0']))
 
 
 # noinspection PyTupleAssignmentBalance
@@ -49,7 +49,7 @@ class LightCurvePRHGenerator:
         self.sf_fit_pars = None
         self.sf_fit_func = None
         self.seed = seed
-        self.h5_dataset = None
+        self.h5_file = None
 
         if self.sf_model not in {'power_law', 'exponential'}:
             raise ValueError(f'Invalid structure function model {self.sf_model}, available '
@@ -65,7 +65,7 @@ class LightCurvePRHGenerator:
         self.qso_lc_vals = self.qso_df[f'mag_{self.image}'].values
         self.qso_lc_errs = self.qso_df[f'magerr_{self.image}'].values
 
-    def estimateStructureFunction(self) -> None:
+    def evaluateStructureFunction(self) -> None:
         tau_matrix = self.computeLagsMatrix(self.t)
         sf_matrix = self.computeSFPointEstimatesMatrix(self.qso_lc_vals, self.qso_lc_errs)
         tau_v_sorted = [[my_tau, my_sf_val] for my_tau, my_sf_val in sorted(zip(tau_matrix.ravel(),
@@ -78,7 +78,7 @@ class LightCurvePRHGenerator:
         self.sf_lags = tau_bin_means
         self.sf_vals = sf_bin_means
 
-    def fitSFModel(self) -> None:
+    def fitStructureFunctionModel(self) -> None:
[…]

### 2020-07-23 paper/article_2.tex
update struttura paper

@@
 Improving the precision in the $H_0$ evaluation will help in finding the reason of this big discrepancy, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{weinberg_euclid}.
+\\
+...
+\\
+The paper is organised as follows.
+In Sec.
+2 we describe the Monte Carlo simulations used in our work to estimate the time delays between light curves.
+In Sec.
+3 we describe the statistical methods to derive time delays and the corresponding errors, where we will also introduce deep gaussian processes.
+In Sec.
+4 we will show the performances of the previous methods applied both on our Monte Carlo simulations and on the real data from COSMOGRAIL \cite{cosmograil}.
+In Sec.
+5 we will discuss about the color variability of quasars.
+In sec.
+6 we will describe our proposed method to estimate the time delay for not resolved lensed quasars by using the color information.
+And finally, in Sec.
+7 we will show the results of the proposed method.
 
@@
 
-\section{Methods}
+\section{Monte Carlo Simulations of Quasars Light Curves}
 
-Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas. 
-\begin{itemize}
-\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
-\item Curabitur feugiat
-\item turpis sed auctor facilisis
-\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
-\item Fusce fermentum, mi sit amet euismod rutrum
-\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
-\item Pellentesque bibendum pretium aliquet
-\end{itemize}
-\blindtext % Dummy text
-
+Parte che potrebbe scrivere Luca Paganin?
+\\
 Text requiring further explanation\footnote{Example footnote}.
@@
 
-\section{Results}
+\section{Statistical Methods to Derive the Time Delay}
 
-\begin{table}
-\caption{Example table}
-\centering
-\begin{tabular}{llr}
-\toprule
-\multicolumn{2}{c}{Name} \\
-\cmidrule(r){1-2}
-First name & Last Name & Grade \\
-\midrule
-John & Doe & $7.5$ \\
-Richard & Miles & $2$ \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\blindtext % Dummy text
-
-\begin{equation}
-\label{eq:emc}
-e = mc^2
-\end{equation}
-
-\blindtext % Dummy text
+Parte che potrebbe scrivere Luca Biggio?
+\\
+(se possibile io metterei anche i deep gaussian processes)
 
@@
 
-\section{Discussion}
+\section{Time Delay Estimations}
+
+ Qui mostrerei i risultati dei metodi applicati sia al Monte Carlo che ai dati di Cosmograil: per il Monte Carlo mostrerei quel bel grafico che ha fatto Luca con il vero DeltaT e quello stimato.
+Per i dati veri farei una tabella tipo
+ \begin{table}[]
+ 	\begin{tabular}{c|c|c}
+ 	Quasar	& $\Delta$t from other searches &  Our $\Delta$t  \\
+ 		&  &    \\
+ 		&  &  \\
+ 		&  & 
+ 	\end{tabular}
+ \end{table}
+Inoltre, parlerei di: 
+\begin{enumerate}
+	\item Come varia la stima di $\Delta$t al variare del campionamento dei gaussian processes
+	\item Commenti sulla stima di $\Delta$t con il metodo standard e con i deep gaussian processes (qui bisogna vedere i risultati)
+\end{enumerate}
+
+\section{Quasars Color Variability}
+Parte che posso scrivere io (Alba).
+
+\section{Method to estimate $\Delta$t in non resolved lensed quasars}
+Parte che posso scrivere io (Alba).
+\\
+Reminder: cita anche il paper del microlensing che in alcuni casi può far variare il colore
+
+\section{Time Delay Estimation from Non Resolved Multiple Images}
+
+Qui mettiamo i risultati.
 
@@
 A statement requiring citation \cite{Figueredo:2009dg}.
-\blindtext % Dummy text
-
-\subsection{Subsection Two}
-
-\blindtext % Dummy text
 

### 2020-07-23 paper/article_2.tex
update struttura paper

@@
  Qui mostrerei i risultati dei metodi applicati sia al Monte Carlo che ai dati di Cosmograil: per il Monte Carlo mostrerei quel bel grafico che ha fatto Luca con il vero DeltaT e quello stimato.
-Per i dati veri farei una tabella tipo
- \begin{table}[]
+Per i dati veri farei una tabella tipo:
+ \begin{table}[h]
  	\begin{tabular}{c|c|c}
@@
  \end{table}
+\\
+Sarebbe fantastico non limitarsi solo ai quasars di cosmograil ma guardare anche https://research.ast.cam.ac.uk/lensedquasars/index.html
+\\
 Inoltre, parlerei di: 
@@
 Parte che posso scrivere io (Alba).
+\\
+The study of quasars light curves has long been seen as a way to understand the structure of the central engine of active galactic nuclei.
+leggi https://academic.oup.com/mnras/article/344/2/492/1124264 per i modelli che spiegano la variabilità dei colori.
 

### 2020-07-23 paper/article_2.tex
update struttura paper
updated TeX documents organization and adopted Mendeley-generated bibliography

@@
 
-\usepackage{blindtext} % Package to generate dummy text throughout this template
+\usepackage{preamble}
 
-\usepackage[sc]{mathpazo} % Use the Palatino font
-\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
+%-------------------
+%	BIBLIOGRAPHY
+%-------------------
 
-\usepackage[english]{babel} % Language hyphenation and typographical rules
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-
-\usepackage{enumitem} % Customized lists
-\setlist[itemize]{noitemsep} % Make itemize lists more compact
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-
-\usepackage{titling} % Customizing the title section
-
-\usepackage{hyperref} % For hyperlinks in the PDF
+\usepackage[style=numeric-comp,useprefix,hyperref,backend=biber,sorting=none,maxbibnames=10]{biblatex}
+\addbibresource{bib_by_hand.bib}
+\addbibresource{QSO_timedelay.bib}
 
@@
 \\
-In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars \cite{h0licow_I} and, gravitational waves [cita].
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars \cite{H0licow_I} and, gravitational waves [cita].
 \\
-The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurements, which has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration \cite{h0licow_XIII}, whose measured value is $H_0=73.3^{+1.7}_{-1.8}$ km s$^{-1}$ Mpc$^{-1}$, in agreement with the distance ladder results and, together with them, with a $5.3 \, \sigma$ tension with the Planck analysis assuming flat $\Lambda$CDM.
+The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurements, which has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration \cite{H0licow_XIII}, whose measured value is $H_0=73.3^{+1.7}_{-1.8}$ km s$^{-1}$ Mpc$^{-1}$, in agreement with the distance ladder results and, together with them, with a $5.3 \, \sigma$ tension with the Planck analysis assuming flat $\Lambda$CDM.
 \\
-In this paper we will focus on the gravitational lensing: firstly suggested by Refsdal \cite{refsdal}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
+In this paper we will focus on the gravitational lensing: firstly suggested by Refsdal \cite{Refsdal1964}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
 This method depends on the matter distribution in the source light trajectory, namely the lensing object (such as a galaxy) and objects along the line of sight, and it has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
@@
 This method requires a long photometric monitoring of the multiple images of the source, of the order of years, and a good temporal sampling, to be able to observe the photometric variations of the source.
-In this regard, the COSMOGRAIL collaboration has been monitoring 18 strongly lensed quasars since 2004 \cite{cosmograil} with 1-2 m size telescopes.
-And the H0LiCOW collaboration has used part of these data to evaluate $H_0$ with a precision of $2.4\%$ \cite{h0licow_XIII}.
+In this regard, the COSMOGRAIL \cite{Cosmograil2020} collaboration has been monitoring 18 strongly lensed quasars since 2004  with 1-2 m size telescopes.
+And the H0LiCOW collaboration has used part of these data to evaluate $H_0$ with a precision of $2.4\%$ \cite{H0licow_XIII}.
 \\
-Improving the precision in the $H_0$ evaluation will help in finding the reason of this big discrepancy, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{weinberg_euclid}.
+Improving the precision in the $H_0$ evaluation will help in finding the reason of this big discrepancy, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{Weinberg2013}.
 \\
@@
 In Sec.
-4 we will show the performances of the previous methods applied both on our Monte Carlo simulations and on the real data from COSMOGRAIL \cite{cosmograil}.
+4 we will show the performances of the previous methods applied both on our Monte Carlo simulations and on the real data from COSMOGRAIL \cite{Cosmograil2020}.
 In Sec.
@@
 Qui mettiamo i risultati.
+\\
+Usare curve Monte Carlo da GERLUMPH?
 
@@
 %----------------------------------------------------------------------------------------
+
+\addcontentsline{toc}{chapter}{\bibname}
+\nocite{*}
+\onecolumn{\printbibliography}
+
+\begin{comment}
 
@@
 
+\end{comment}
+
 %----------------------------------------------------------------------------------------

### 2020-07-23 paper/update_bib.py

@@ -1,0 +1,12 @@
+from pathlib import Path
+import shutil
+import os
+
+srcdir = Path(os.getenv("HOME")) / "OnlyLocalFiles/Mendeley_bibfiles"
+srcfile = srcdir / "QSO time delay.bib"
+
+dstfile = Path('.') / "QSO_timedelay.bib"
+
+if dstfile.exists():
+    dstfile.unlink()
+shutil.copy(srcfile, dstfile)

### 2020-07-24 paper/article_2.tex
Update article_2.tex
modifiche minimali a abstract e introduzione

@@
 \begin{abstract}
-\noindent The recent tensions on the measured value of the Hubble constant between CMB and astrophisical observations, has triggered the need of new methods for its determination.
-In view of this, an effort has been done by H0LiCoW to use the gravitational lensing of quasars as a probe for H0.
-This type of measurement requires a long term monitoring of lensed quasars (of the order of years).
-Since big telescopes have to deal with many observational requests, it is difficult to have a constant monitoring over the years, therefore this task can be achieved more easily by small/medium size telescopes.
-However, the number of lensed quasars with multiple images that can be resolved by these telescopes drops drastically.
-Here we present a method to deal with \textbf{non resolved} lensed quasars.
+\noindent The recent tensions on the measured value of the Hubble parameter $H_0$ between cosmic microwave background radiation and astrophysical observations have triggered the need of new methods for its determination.
+In view of this, an effort has been done by the H0LiCoW collaboration to use the gravitational lensing of quasars as a probe for $H_0$ through the measurement of the time delay between the multiple images of the lensed quasar.
+Measurements of the time delay require a long term monitoring of lensed quasars, of the order of years.
+Since big telescopes have to deal with many observational requests, it is difficult to have a regular monitoring campaign over the years, therefore this task can be achieved more easily by small/medium sized telescopes.
+However, the number of lensed quasars with multiple images that can be angularly resolved by these telescopes drops drastically.
+Here we present a method to deal with \textbf{non fully resolved} lensed quasars.
 This method has also the advantage of being less dependent on the microlensing effect of the lens galaxy.
@@
 \section{Introduction}
-In the last years, the precision of the Planck experiment [cita], whose task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model ($\Lambda$CDM) which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
-In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the current expansion rate of the Universe, $H_0$, whose inference strongly depends on the assumed cosmological model.
+In the last years, the precision of the Planck experiment [cita], whose main task was to analyse the Cosmic Microwave Background (CMB) anisotropies, has allowed to fully test our standard cosmological model ($\Lambda$CDM) which assumes the existence of Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
+In particular, in addition to the minimal 6 parameters describing $\Lambda$CDM, the CMB anisotropies allow to indirectly constrain other parameters, such as the Hubble parameter $H_0$, which represents the current expansion rate of the Universe.
+The inference of $H_0$ strongly depends on the assumed cosmological model.
 For example, relaxing the spatial flatness hypothesis of our Universe or the constant equation of state for the dark energy, would impact the $H_0$ estimation.
 \\
-In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitational lensed quasars \cite{H0licow_I} and, gravitational waves [cita].
+In parallel, the are other independent methods to measure $H_0$, such as the distance ladder [cita], water masers [cita], the time delay between multiple images of gravitationally lensed quasars \cite{H0licow_I} and, gravitational waves [cita].
 \\
-The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurements, which has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration \cite{H0licow_XIII}, whose measured value is $H_0=73.3^{+1.7}_{-1.8}$ km s$^{-1}$ Mpc$^{-1}$, in agreement with the distance ladder results and, together with them, with a $5.3 \, \sigma$ tension with the Planck analysis assuming flat $\Lambda$CDM.
+The highest precision reached by Plank has however shown a tension in the value of $H_0$ with respect to the distance ladder measurements, which has been further enhanced by the recent gravitational lensing results from the H0LiCOW collaboration \cite{H0licow_XIII}, whose measured value is $H_0=73.3^{+1.7}_{-1.8}$ km s$^{-1}$ Mpc$^{-1}$, in agreement with the distance ladder results and, together with them, with a $5.3 \, \sigma$ tension with the Planck analysis, assuming the flat $\Lambda$CDM model.
 \\
-In this paper we will focus on the gravitational lensing: firstly suggested by Refsdal \cite{Refsdal1964}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
-This method depends on the matter distribution in the source light trajectory, namely the lensing object (such as a galaxy) and objects along the line of sight, and it has a weaker dependence on the cosmological parameters if compared to the CMB analyses.
+In this paper we will focus on the gravitational lensing method: firstly suggested by Refsdal \cite{Refsdal1964}, this method directly relates the time delays between multiple images of the same source produced by a lensing object with $H_0$ in the form $\Delta_T \propto 1/H_0$.
+This method depends on the matter distribution in the light trajectory from the source, namely the lensing object (such as a galaxy or a cluster of galaxies) and objects along the line of sight, and it has a weaker dependence on the cosmological model if compared to the CMB analyses.
 In particular, it depends on the matter density $\Omega_m$, the dark energy density $\Omega_\Lambda$, the curvature parameter $\Omega_k$ and the dark energy equation of state $\omega$ [cita].
 \\
-This method requires a long photometric monitoring of the multiple images of the source, of the order of years, and a good temporal sampling, to be able to observe the photometric variations of the source.
+This method requires a long photometric monitoring of the multiple images of the source, of the order of years, and a proper temporal sampling, to be able to observe the photometric variations of the source.
 In this regard, the COSMOGRAIL \cite{Cosmograil2020} collaboration has been monitoring 18 strongly lensed quasars since 2004  with 1-2 m size telescopes.
@@
 \\
-Improving the precision in the $H_0$ evaluation will help in finding the reason of this big discrepancy, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{Weinberg2013}.
+Improving the precision in the $H_0$ evaluation will help in finding the reason of the big discrepancy between the measured values of $H_0$, and, it would also have a big impact in the results of the next cosmological surveys, up to a 40$\%$ improvement if $H_0$ is independently known with $1\%$ precision \cite{Weinberg2013}.
 \\
-...
+In this paper we propose a novel method to improve the statistics of available time delay measurements of lensed quasars to be used for the determination of $H_0$.
+In fact, as said, monitoring campaigns are often conducted by small-medium sized telescopes, that have a reduced angular resolution with respect to big telescopes, hence the available sample of light curves of optically fully-resolved multiple images of lensed quasars is limited.
+In fact, several gravitationally-lensed quasars have angular separation of 1.5 arcsec or smaller between the multiple images (ref.
+CASTLE database), at the edge of the resolution of many small telescopes.
+However, we propose a method such that, even when the multiple images of the source cannot be fully resolved by the telescope, still a time delay can be retrieved if data are taken in at least two different photometric filters, exploiting a feature of the variability of the luminosity of quasars, which is not uniform among various frequency bands.
+The time delay is then extracted not from (at least) two light curves of different resolved images, rather from (at least) two light curves of not-resolved images in different photometric filters.
 \\

### 2020-08-07 paper/article_2.tex
update paper

@@
 \documentclass[twoside,twocolumn]{article}
+
+\usepackage{xcolor}
 
@@
 Usare curve Monte Carlo da GERLUMPH?
+\\
+\textcolor{red}{Un'obiezione che possono fare è che nel caso delle lenti non risolte, non si riesce comunque ad osservare bene nel dettaglio la forma della galassia lente (nemmeno con il telescopio Hubble), e questo porta ad un errore maggiore nella stima di H0.
+Noi dobbiamo evidenziare il fatto che nuovi telescopi ad altissima risoluzione, quali Euclid (altro collegamento per Luca) e FORSE il telescopio James Webb.}
 

### 2020-08-07 paper/article_2.tex
update paper

@@
 \textcolor{red}{Un'obiezione che possono fare è che nel caso delle lenti non risolte, non si riesce comunque ad osservare bene nel dettaglio la forma della galassia lente (nemmeno con il telescopio Hubble), e questo porta ad un errore maggiore nella stima di H0.
-Noi dobbiamo evidenziare il fatto che nuovi telescopi ad altissima risoluzione, quali Euclid (altro collegamento per Luca) e FORSE il telescopio James Webb.}
+Noi dobbiamo evidenziare il fatto che nuovi telescopi ad altissima risoluzione, quali Euclid (altro collegamento per Luca), FORSE il telescopio James Webb, e infine LSST.}
 

### 2020-11-21 paper/aa-package/paper.tex
new version of the paper for A&A journal (voliamo alti)

%                                                                 aa.dem
% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\begin{document}

   \title{Time delay estimation in non-resolved lensed quasars}

   \author{G. Wuchterl
          \inst{1}
          \and
          C. Ptolemy\inst{2}\fnmsep\thanks{Just to show the usage
          of the elements in the author field}
          }

   \institute{Institute for Astronomy (IfA), University of Vienna,
              T\"urkenschanzstrasse 17, A-1180 Vienna\\
              \email{wuchterl@amok.ast.univie.ac.at}
         \and
             University of Alexandria, Department of Geography, ...\\
             \email{c.ptolemy@hipparch.uheaven.space}
             \thanks{The university of heaven temporarily does not
                     accept e-mails}
             }

   \date{Received September 15, 1996; accepted March 16, 1997}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {
   	The analysis of Cosmic Microwave Background anisotropies and the cosmic distance ladder observations have resulted in a tension on the estimated value of the Hubble parameter $H_0$. This has triggered the need for new methods for its determination. To this extent, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ based on time delay measurements of six gravitational lensed quasars. Such value is compatible with the distance ladder observations, increasing the tension at the $5.3 \sigma$ level.
   	}
  % aims heading (mandatory)
   { 
   	The fractional error on $H_0$, obtained with the gravitational lensing method, decreases linearly with the number of lensed systems analysed. Moreover, such method is highly affected by the microlensing phenomenon, which can produce changes in the actual time delay of the order of days. To account for these two facts, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring of the lensed source, of the order of years. This cannot be achieved with big telescopes, due to the huge amount of observational requests they have to fulfill. On the other hand, small/medium-sized telescopes are present in a much larger number and are often characterized by more versatile observational programs. However, the limited resolution capabilities of such instruments and their often not privileged geographical location prevent them from providing well-separated images of the same lensed source. 
   	\\
   	In this work, we present a novel approach to estimate the time-delay in unresolved lensed quasar systems. Our method is further motivated by recent developments in discovering more unresolved strongly-lensed QSO systems.
	}
  % methods heading (mandatory)
   {Our method uses ...}
  % results heading (mandatory)
   {...}
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{giant planet formation --
                $\kappa$-mechanism --
                stability of gas spheres
               }

   \maketitle
%
%-------------------------------------------------------------------

\section{Introduction}

   In the \emph{nucleated instability\/} (also called core
   instability) hypothesis of giant planet
   formation, a critical mass for static core  envelope
   protoplanets has been found. \citet{mizuno} determined
   the critical mass of the core to be about $12 \,M_\oplus$
   ($M_\oplus=5.975 \times 10^{27}\,\mathrm{g}$ is the Earth mass), which
   is independent of the outer boundary
   conditions and therefore independent of the location in the
   solar nebula. This critical value for the core mass corresponds
   closely to the cores of today's giant planets.

   Although no hydrodynamical study has been available many workers
   conjectured that a collapse or rapid contraction will ensue
   after accumulating the critical mass. The main motivation for
   this article
   is to investigate the stability of the static envelope at the
   critical mass. With this aim the local, linear stability of static
   radiative gas  spheres is investigated on the basis of Baker's
   (\citeyear{baker}) standard one-zone model.

   Phenomena similar to the ones described above for giant planet
   formation have been found in hydrodynamical models concerning
   star formation where protostellar cores explode
   (Tscharnuter \citeyear{tscharnuter}, Balluch \citeyear{balluch}),
   whereas earlier studies found quasi-steady collapse flows. The
   similarities in the (micro)physics, i.e., constitutive relations of
   protostellar cores and protogiant planets serve as a further
   motivation for this study.

%--------------------------------------------------------------------
\section{Baker's standard one-zone model}

%-------------------------------------- Two column figure (place early!)
   \begin{figure*}
   \centering
   %%%\includegraphics{empty.eps}
   %%%\includegraphics{empty.eps}
   %%%\includegraphics{empty.eps}
   \caption{Adiabatic exponent $\Gamma_1$.
               $\Gamma_1$ is plotted as a function of
               $\lg$ internal energy $\mathrm{[erg\,g^{-1}]}$ and $\lg$
               density $\mathrm{[g\,cm^{-3}]}$.}
              \label{FigGam}%
    \end{figure*}
%
   In this section the one-zone model of \citet{baker},
   originally used to study the Cephe{\"{\i}}d pulsation mechanism, will
   be briefly reviewed. The resulting stability criteria will be
   rewritten in terms of local state variables, local timescales and
   constitutive relations.

   \citet{baker} investigates the stability of thin layers in
   self-gravitating,
   spherical gas clouds with the following properties:
   \begin{itemize}
      \item hydrostatic equilibrium,
      \item thermal equilibrium,
      \item energy transport by grey radiation diffusion.
   \end{itemize}
   For the one-zone-model Baker obtains necessary conditions
   for dynamical, secular and vibrational (or pulsational)
   stability (Eqs.\ (34a,\,b,\,c) in Baker \citeyear{baker}). Using Baker's
   notation:
   \[
      \begin{array}{lp{0.8\linewidth}}
         M_{r}  & mass internal to the radius $r$     \\
         m               & mass of the zone                    \\
         r_0             & unperturbed zone radius             \\
         \rho_0          & unperturbed density in the zone     \\
         T_0             & unperturbed temperature in the zone \\
         L_{r0}          & unperturbed luminosity              \\
         E_{\mathrm{th}} & thermal energy of the zone
      \end{array}
   \]
\noindent
   and with the definitions of the \emph{local cooling time\/}
   (see Fig.~\ref{FigGam})
   \begin{equation}
      \tau_{\mathrm{co}} = \frac{E_{\mathrm{th}}}{L_{r0}} \,,
   \end{equation}
   and the \emph{local free-fall time}
   \begin{equation}
      \tau_{\mathrm{ff}} =
         \sqrt{ \frac{3 \pi}{32 G} \frac{4\pi r_0^3}{3 M_{\mathrm{r}}}
}\,,
   \end{equation}
   Baker's $K$ and $\sigma_0$ have the following form:
   \begin{eqnarray}
      \sigma_0 & = & \frac{\pi}{\sqrt{8}}
                     \frac{1}{ \tau_{\mathrm{ff}}} \\
      K        & = & \frac{\sqrt{32}}{\pi} \frac{1}{\delta}
                        \frac{ \tau_{\mathrm{ff}} }
                             { \tau_{\mathrm{co}} }\,;
   \end{eqnarray}
   where $ E_{\mathrm{th}} \approx m (P_0/{\rho_0})$ has been used and
   \begin{equation}
   \begin{array}{l}
      \delta = - \left(
                    \frac{ \partial \ln \rho }{ \partial \ln T }
                 \right)_P \\
      e=mc^2
   \end{array}
   \end{equation}
   is a thermodynamical quantity which is of order $1$ and equal to $1$
   for nonreacting mixtures of classical perfect gases. The physical
   meaning of $ \sigma_0 $ and $K$ is clearly visible in the equations
   above. $\sigma_0$ represents a frequency of the order one per
   free-fall time. $K$ is proportional to the ratio of the free-fall
   time and the cooling time. Substituting into Baker's criteria, using
   thermodynamic identities and definitions of thermodynamic quantities,
   \begin{displaymath}
      \Gamma_1      = \left( \frac{ \partial \ln P}{ \partial\ln \rho}
                           \right)_{S}    \, , \;
      \chi^{}_\rho  = \left( \frac{ \partial \ln P}{ \partial\ln \rho}
                           \right)_{T}    \, , \;
      \kappa^{}_{P} = \left( \frac{ \partial \ln \kappa}{ \partial\ln P}
                           \right)_{T}
   \end{displaymath}
   \begin{displaymath}
      \nabla_{\mathrm{ad}} = \left( \frac{ \partial \ln T}
                             { \partial\ln P} \right)_{S} \, , \;
      \chi^{}_T       = \left( \frac{ \partial \ln P}
                             { \partial\ln T} \right)_{\rho} \, , \;
      \kappa^{}_{T}   = \left( \frac{ \partial \ln \kappa}
                             { \partial\ln T} \right)_{T}
   \end{displaymath}
   one obtains, after some pages of algebra, the conditions for
   \emph{stability\/} given
   below:
   \begin{eqnarray}
      \frac{\pi^2}{8} \frac{1}{\tau_{\mathrm{ff}}^2}
                ( 3 \Gamma_1 - 4 )
         & > & 0 \label{ZSDynSta} \\
      \frac{\pi^2}{\tau_{\mathrm{co}}
                   \tau_{\mathrm{ff}}^2}
                   \Gamma_1 \nabla_{\mathrm{ad}}
                   \left[ \frac{ 1- 3/4 \chi^{}_\rho }{ \chi^{}_T }
                          ( \kappa^{}_T - 4 )
                        + \kappa^{}_P + 1
                   \right]
        & > & 0 \label{ZSSecSta} \\
     \frac{\pi^2}{4} \frac{3}{\tau_{ \mathrm{co} }
                              \tau_{ \mathrm{ff} }^2
                             }
         \Gamma_1^2 \, \nabla_{\mathrm{ad}} \left[
                                   4 \nabla_{\mathrm{ad}}
                                   - ( \nabla_{\mathrm{ad}} \kappa^{}_T
                                     + \kappa^{}_P
                                     )
                                   - \frac{4}{3 \Gamma_1}
                                \right]
        & > & 0   \label{ZSVibSta}
   \end{eqnarray}
%
   For a physical discussion of the stability criteria see \citet{baker} or \citet{cox}.

   We observe that these criteria for dynamical, secular and
   vibrational stability, respectively, can be factorized into
   \begin{enumerate}
      \item a factor containing local timescales only,
      \item a factor containing only constitutive relations and
         their derivatives.
   \end{enumerate}
   The first factors, depending on only timescales, are positive
   by definition. The signs of the left hand sides of the
   inequalities~(\ref{ZSDynSta}), (\ref{ZSSecSta}) and (\ref{ZSVibSta})
   therefore depend exclusively on the second factors containing
   the constitutive relations. Since they depend only
   on state variables, the stability criteria themselves are \emph{
   functions of the thermodynamic state in the local zone}. The
   one-zone stability can therefore be determined
   from a simple equation of state, given for example, as a function
   of density and
   temperature. Once the microphysics, i.e.\ the thermodynamics
   and opacities (see Table~\ref{KapSou}), are specified (in practice
   by specifying a chemical composition) the one-zone stability can
   be inferred if the thermodynamic state is specified.
   The zone -- or in
   other words the layer -- will be stable or unstable in
   whatever object it is imbedded as long as it satisfies the
   one-zone-model assumptions. Only the specific growth rates
   (depending upon the time scales) will be different for layers
   in different objects.

%--------------------------------------------------- One column table
   \begin{table}
      \caption[]{Opacity sources.}
         \label{KapSou}
     $$ 
         \begin{array}{p{0.5\linewidth}l}
            \hline
            \noalign{\smallskip}
            Source      &  T / {[\mathrm{K}]} \\
            \noalign{\smallskip}
            \hline
            \noalign{\smallskip}
            Yorke 1979, Yorke 1980a & \leq 1700^{\mathrm{a}}     \\
%           Yorke 1979, Yorke 1980a & \leq 1700             \\
            Kr\"ugel 1971           & 1700 \leq T \leq 5000 \\
            Cox \& Stewart 1969     & 5000 \leq             \\
            \noalign{\smallskip}
            \hline
         \end{array}
     $$ 
   \end{table}
%
   We will now write down the sign (and therefore stability)
   determining parts of the left-hand sides of the inequalities
   (\ref{ZSDynSta}), (\ref{ZSSecSta}) and (\ref{ZSVibSta}) and thereby
   obtain \emph{stability equations of state}.

   The sign determining part of inequality~(\ref{ZSDynSta}) is
   $3\Gamma_1 - 4$ and it reduces to the
   criterion for dynamical stability
   \begin{equation}
     \Gamma_1 > \frac{4}{3}\,\cdot
   \end{equation}
   Stability of the thermodynamical equilibrium demands
   \begin{equation}
      \chi^{}_\rho > 0, \;\;  c_v > 0\, ,
   \end{equation}
   and
   \begin{equation}
      \chi^{}_T > 0
   \end{equation}
   holds for a wide range of physical situations.
   With
   \begin{eqnarray}
      \Gamma_3 - 1 = \frac{P}{\rho T} \frac{\chi^{}_T}{c_v}&>&0\\
      \Gamma_1     = \chi_\rho^{} + \chi_T^{} (\Gamma_3 -1)&>&0\\
      \nabla_{\mathrm{ad}}  = \frac{\Gamma_3 - 1}{\Gamma_1}         &>&0
   \end{eqnarray}
   we find the sign determining terms in inequalities~(\ref{ZSSecSta})
   and (\ref{ZSVibSta}) respectively and obtain the following form
   of the criteria for dynamical, secular and vibrational
   \emph{stability}, respectively:
   \begin{eqnarray}
      3 \Gamma_1 - 4 =: S_{\mathrm{dyn}}      > & 0 & \label{DynSta}  \\
%
      \frac{ 1- 3/4 \chi^{}_\rho }{ \chi^{}_T } ( \kappa^{}_T - 4 )
         + \kappa^{}_P + 1 =: S_{\mathrm{sec}} > & 0 & \label{SecSta} \\
%
      4 \nabla_{\mathrm{ad}} - (\nabla_{\mathrm{ad}} \kappa^{}_T
                             + \kappa^{}_P)
                             - \frac{4}{3 \Gamma_1} =: S_{\mathrm{vib}}
                                      > & 0\,.& \label{VibSta}
   \end{eqnarray}
   The constitutive relations are to be evaluated for the
   unperturbed thermodynamic state (say $(\rho_0, T_0)$) of the zone.
   We see that the one-zone stability of the layer depends only on
   the constitutive relations $\Gamma_1$,
   $\nabla_{\mathrm{ad}}$, $\chi_T^{},\,\chi_\rho^{}$,
   $\kappa_P^{},\,\kappa_T^{}$.
   These depend only on the unperturbed
   thermodynamical state of the layer. Therefore the above relations
   define the one-zone-stability equations of state
   $S_{\mathrm{dyn}},\,S_{\mathrm{sec}}$
   and $S_{\mathrm{vib}}$. See Fig.~\ref{FigVibStab} for a picture of
   $S_{\mathrm{vib}}$. Regions of secular instability are
   listed in Table~1.

%
%                                                One column figure
%----------------------------------------------------------------- 
   \begin{figure}
   \centering
   %%%\includegraphics[width=3cm]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%-----------------------------------------------------------------

\section{Conclusions}

   \begin{enumerate}
      \item The conditions for the stability of static, radiative
         layers in gas spheres, as described by Baker's (\citeyear{baker})
         standard one-zone model, can be expressed as stability
         equations of state. These stability equations of state depend
         only on the local thermodynamic state of the layer.
      \item If the constitutive relations -- equations of state and
         Rosseland mean opacities -- are specified, the stability
         equations of state can be evaluated without specifying
         properties of the layer.
      \item For solar composition gas the $\kappa$-mechanism is
         working in the regions of the ice and dust features
         in the opacities, the $\mathrm{H}_2$ dissociation and the
         combined H, first He ionization zone, as
         indicated by vibrational instability. These regions
         of instability are much larger in extent and degree of
         instability than the second He ionization zone
         that drives the Cephe{\"\i}d pulsations.
   \end{enumerate}

\begin{acknowledgements}
      Part of this work was supported by the German
      \emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project
      number Ts~17/2--1.
\end{acknowledgements}

% WARNING
%-------------------------------------------------------------------
% Please note that we have included the references to the file aa.dem in
% order to compile it, but we ask you to:
%
% - use BibTeX with the regular commands:
%   \bibliographystyle{aa} % style aa.bst
%   \bibliography{Yourfile} % your references Yourfile.bib
%
% - join the .bib files when you upload your source files
%-------------------------------------------------------------------

\begin{thebibliography}{}

  \bibitem[Baker(1966)]{baker} Baker, N. 1966,
      in Stellar Evolution,
      ed.\ R. F. Stein,\& A. G. W. Cameron
      (Plenum, New York) 333

   \bibitem[Balluch(1988)]{balluch} Balluch, M. 1988,
      A\&A, 200, 58

   \bibitem[Cox(1980)]{cox} Cox, J. P. 1980,
      Theory of Stellar Pulsation
      (Princeton University Press, Princeton) 165

   \bibitem[Cox(1969)]{cox69} Cox, A. N.,\& Stewart, J. N. 1969,
      Academia Nauk, Scientific Information 15, 1

   \bibitem[Mizuno(1980)]{mizuno} Mizuno H. 1980,
      Prog. Theor. Phys., 64, 544

   \bibitem[Tscharnuter(1987)]{tscharnuter} Tscharnuter W. M. 1987,
      A\&A, 188, 55

   \bibitem[Terlevich(1992)]{terlevich} Terlevich, R. 1992, in ASP Conf. Ser. 31, 
      Relationships between Active Galactic Nuclei and Starburst Galaxies, 
      ed. A. V. Filippenko, 13

   \bibitem[Yorke(1980a)]{yorke80a} Yorke, H. W. 1980a,
      A\&A, 86, 286

   \bibitem[Zheng(1997)]{zheng} Zheng, W., Davidsen, A. F., Tytler, D. \& Kriss, G. A.
      1997, preprint
\end{thebibliography}

\end{document}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Example below of non-structurated natbib references  
To use the v8.3 macros with this form of composition of bibliography, 
the option "bibyear" should be added to the command line 
"\documentclass[bibyear]{aa}".
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{}

  \bibitem[1966]{baker} Baker, N. 1966,
      in Stellar Evolution,
      ed.\ R. F. Stein,\& A. G. W. Cameron
      (Plenum, New York) 333

   \bibitem[1988]{balluch} Balluch, M. 1988,
      A\&A, 200, 58

   \bibitem[1980]{cox} Cox, J. P. 1980,
      Theory of Stellar Pulsation
      (Princeton University Press, Princeton) 165

   \bibitem[1969]{cox69} Cox, A. N.,\& Stewart, J. N. 1969,
      Academia Nauk, Scientific Information 15, 1

   \bibitem[1980]{mizuno} Mizuno H. 1980,
      Prog. Theor. Phys., 64, 544

   \bibitem[1987]{tscharnuter} Tscharnuter W. M. 1987,
      A\&A, 188, 55

   \bibitem[1992]{terlevich} Terlevich, R. 1992, in ASP Conf. Ser. 31, 
      Relationships between Active Galactic Nuclei and Starburst Galaxies, 
      ed. A. V. Filippenko, 13

   \bibitem[1980a]{yorke80a} Yorke, H. W. 1980a,
      A\&A, 86, 286

   \bibitem[1997]{zheng} Zheng, W., Davidsen, A. F., Tytler, D. \& Kriss, G. A.
      1997, preprint
\end{thebibliography}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Examples for figures using graphicx
A guide "Using Imported Graphics in LaTeX2e"  (Keith Reckdahl)
is available on a lot of LaTeX public servers or ctan mirrors.
The file is : epslatex.pdf 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------------------
%                 A figure as large as the width of the column
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[width=\hsize]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%-------------------------------------------------------------
%                                    One column rotated figure
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[angle=-90,width=3cm]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%-------------------------------------------------------------
%                        Figure with caption on the right side 
%-------------------------------------------------------------
   \begin{figure}
   \sidecaption
   \includegraphics[width=3cm]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%-------------------------------------------------------------
%                                Figure with a new BoundingBox 
%-------------------------------------------------------------
   \begin{figure}
   \centering
   \includegraphics[bb=10 20 100 300,width=3cm,clip]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%-------------------------------------------------------------
%                                      The "resizebox" command 
%-------------------------------------------------------------
   \begin{figure}
   \resizebox{\hsize}{!}
            {\includegraphics[bb=10 20 100 300,clip]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure}
%
%-------------------------------------------------------------
%                                             Two column Figure 
%-------------------------------------------------------------
   \begin{figure*}
   \resizebox{\hsize}{!}
            {\includegraphics[bb=10 20 100 300,clip]{empty.eps}
      \caption{Vibrational stability equation of state
               $S_{\mathrm{vib}}(\lg e, \lg \rho)$.
               $>0$ means vibrational stability.
              }
         \label{FigVibStab}
   \end{figure*}
%
%-------------------------------------------------------------
%                                             Simple A&A Table
%-------------------------------------------------------------
%
\begin{table}
\caption{Nonlinear Model Results}             % title of Table
\label{table:1}      % is used to refer this table in the text
\centering                          % used for centering table
\begin{tabular}{c c c c}        % centered columns (4 columns)
\hline\hline                 % inserts double horizontal lines
HJD & $E$ & Method\#2 & Method\#3 \\    % table heading 
\hline                        % inserts single horizontal line
   1 & 50 & $-837$ & 970 \\      % inserting body of the table
   2 & 47 & 877    & 230 \\
   3 & 31 & 25     & 415 \\
   4 & 35 & 144    & 2356 \\
   5 & 45 & 300    & 556 \\ 
\hline                                   %inserts single line
\end{tabular}
\end{table}
%
%-------------------------------------------------------------
%                                             Two column Table 
%-------------------------------------------------------------
%
\begin{table*}
\caption{Nonlinear Model Results}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c l l l }     % 7 columns 
\hline\hline       
                      % To combine 4 columns into a single one 
HJD & $E$ & Method\#2 & \multicolumn{4}{c}{Method\#3}\\ 
\hline                    
   1 & 50 & $-837$ & 970 & 65 & 67 & 78\\  
   2 & 47 & 877    & 230 & 567& 55 & 78\\
   3 & 31 & 25     & 415 & 567& 55 & 78\\
   4 & 35 & 144    & 2356& 567& 55 & 78 \\
   5 & 45 & 300    & 556 & 567& 55 & 78\\
\hline                  
\end{tabular}
\end{table*}
%
%-------------------------------------------------------------
%                                          Table with notes 
%-------------------------------------------------------------
%
% A single note
\begin{table}
\caption{\label{t7}Spectral types and photometry for stars in the
  region.}
\centering
\begin{tabular}{lccc}
\hline\hline
Star&Spectral type&RA(J2000)&Dec(J2000)\\
\hline
69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
LS~1267~(86) &O8\,V     &09 15 52.787&11.07\\
24.6         &7.58      &1.37 &0.20\\
\hline
LS~1262      &B0\,V     &09 15 05.17&11.17\\
MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\\
LS~1269      &O8.5\,V   &09 15 56.60&10.85\\
\hline
\end{tabular}
\tablefoot{The top panel shows likely members of Pismis~11. The second
panel contains likely members of Alicante~5. The bottom panel
displays stars outside the clusters.}
\end{table}
%
% More notes
%
\begin{table}
\caption{\label{t7}Spectral types and photometry for stars in the
  region.}
\centering
\begin{tabular}{lccc}
\hline\hline
Star&Spectral type&RA(J2000)&Dec(J2000)\\
\hline
69           &B1\,V     &09 15 54.046 & $-$50 00 26.67\\
49           &B0.7\,V   &*09 15 54.570& $-$50 00 03.90\\
LS~1267~(86) &O8\,V     &09 15 52.787&11.07\tablefootmark{a}\\
24.6         &7.58\tablefootmark{1}&1.37\tablefootmark{a}   &0.20\tablefootmark{a}\\
\hline
LS~1262      &B0\,V     &09 15 05.17&11.17\tablefootmark{b}\\
MO 2-119     &B0.5\,V   &09 15 33.7 &11.74\tablefootmark{c}\\
LS~1269      &O8.5\,V   &09 15 56.60&10.85\tablefootmark{d}\\
\hline
\end{tabular}
\tablefoot{The top panel shows likely members of Pismis~11. The second
panel contains likely members of Alicante~5. The bottom panel
displays stars outside the clusters.\\
\tablefoottext{a}{Photometry for MF13, LS~1267 and HD~80077 from
Dupont et al.}
\tablefoottext{b}{Photometry for LS~1262, LS~1269 from
Durand et al.}
\tablefoottext{c}{Photometry for MO2-119 from
Mathieu et al.}
}
\end{table}
%
%-------------------------------------------------------------
%                                       Table with references 
%-------------------------------------------------------------
%
\begin{table*}[h]
 \caption[]{\label{nearbylistaa2}List of nearby SNe used in this work.}
\begin{tabular}{lccc}
 \hline \hline
  SN name &
  Epoch &
 Bands &
  References \\
 &
  (with respect to $B$ maximum) &
 &
 \\ \hline
1981B   & 0 & {\it UBV} & 1\\
1986G   &  $-$3, $-$1, 0, 1, 2 & {\it BV}  & 2\\
1989B   & $-$5, $-$1, 0, 3, 5 & {\it UBVRI}  & 3, 4\\
1990N   & 2, 7 & {\it UBVRI}  & 5\\
1991M   & 3 & {\it VRI}  & 6\\
\hline
\noalign{\smallskip}
\multicolumn{4}{c}{ SNe 91bg-like} \\
\noalign{\smallskip}
\hline
1991bg   & 1, 2 & {\it BVRI}  & 7\\
1999by   & $-$5, $-$4, $-$3, 3, 4, 5 & {\it UBVRI}  & 8\\
\hline
\noalign{\smallskip}
\multicolumn{4}{c}{ SNe 91T-like} \\
\noalign{\smallskip}
\hline
1991T   & $-$3, 0 & {\it UBVRI}  &  9, 10\\
2000cx  & $-$3, $-$2, 0, 1, 5 & {\it UBVRI}  & 11\\ %
\hline
\end{tabular}
\tablebib{(1)~\citet{branch83};
(2) \citet{phillips87}; (3) \citet{barbon90}; (4) \citet{wells94};
(5) \citet{mazzali93}; (6) \citet{gomez98}; (7) \citet{kirshner93};
(8) \citet{patat96}; (9) \citet{salvo01}; (10) \citet{branch03};
(11) \citet{jha99}.
}
\end{table}
%-------------------------------------------------------------
%                      A rotated Two column Table in landscape  
%-------------------------------------------------------------
\begin{sidewaystable*}
\caption{Summary for ISOCAM sources with mid-IR excess 
(YSO candidates).}\label{YSOtable}
\centering
\begin{tabular}{crrlcl} 
\hline\hline             
ISO-L1551 & $F_{6.7}$~[mJy] & $\alpha_{6.7-14.3}$ 
& YSO type$^{d}$ & Status & Comments\\
\hline
  \multicolumn{6}{c}{\it New YSO candidates}\\ % To combine 6 columns into a single one
\hline
  1 & 1.56 $\pm$ 0.47 & --    & Class II$^{c}$ & New & Mid\\
  2 & 0.79:           & 0.97: & Class II ?     & New & \\
  3 & 4.95 $\pm$ 0.68 & 3.18  & Class II / III & New & \\
  5 & 1.44 $\pm$ 0.33 & 1.88  & Class II       & New & \\
\hline
  \multicolumn{6}{c}{\it Previously known YSOs} \\
\hline
  61 & 0.89 $\pm$ 0.58 & 1.77 & Class I & \object{HH 30} & Circumstellar disk\\
  96 & 38.34 $\pm$ 0.71 & 37.5& Class II& MHO 5          & Spectral type\\
\hline
\end{tabular}
\end{sidewaystable*}
%-------------------------------------------------------------
%                      A rotated One column Table in landscape  
%-------------------------------------------------------------
\begin{sidewaystable}
\caption{Summary for ISOCAM sources with mid-IR excess 
(YSO candidates).}\label{YSOtable}
\centering
\begin{tabular}{crrlcl} 
\hline\hline             
ISO-L1551 & $F_{6.7}$~[mJy] & $\alpha_{6.7-14.3}$ 
& YSO type$^{d}$ & Status & Comments\\
\hline
  \multicolumn{6}{c}{\it New YSO candidates}\\ % To combine 6 columns into a single one
\hline
  1 & 1.56 $\pm$ 0.47 & --    & Class II$^{c}$ & New & Mid\\
  2 & 0.79:           & 0.97: & Class II ?     & New & \\
  3 & 4.95 $\pm$ 0.68 & 3.18  & Class II / III & New & \\
  5 & 1.44 $\pm$ 0.33 & 1.88  & Class II       & New & \\
\hline
  \multicolumn{6}{c}{\it Previously known YSOs} \\
\hline
  61 & 0.89 $\pm$ 0.58 & 1.77 & Class I & \object{HH 30} & Circumstellar disk\\
  96 & 38.34 $\pm$ 0.71 & 37.5& Class II& MHO 5          & Spectral type\\
\hline
\end{tabular}
\end{sidewaystable}
%
%-------------------------------------------------------------
%                              Table longer than a single page  
%-------------------------------------------------------------
% All long tables will be placed automatically at the end of the document
%
\longtab{
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86 
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
}
%
%-------------------------------------------------------------
%                              Table longer than a single page
%                                            and in landscape, 
%                    in the preamble, use: \usepackage{lscape}
%-------------------------------------------------------------

% All long tables will be placed automatically at the end of the document
%
\longtab{
\begin{landscape}
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
\end{landscape}
}
%
%-------------------------------------------------------------
%               Appendices have to be placed at the end, after
%                                        \end{thebibliography}
%-------------------------------------------------------------
\end{thebibliography}

\begin{appendix} %First appendix
\section{Background galaxy number counts and shear noise-levels}
Because the optical images used in this analysis...
\begin{figure*}%f1
\includegraphics[width=10.9cm]{1787f23.eps}
\caption{Shown in greyscale is a...}
\label{cl12301}
\end{figure*}

In this case....
\begin{figure*}
\centering
\includegraphics[width=16.4cm,clip]{1787f24.ps}
\caption{Plotted above...}
\label{appfig}
\end{figure*}

Because the optical images...

\section{Title of Second appendix.....} %Second appendix
These studies, however, have faced...
\begin{table}
\caption{Complexes characterisation.}\label{starbursts}
\centering
\begin{tabular}{lccc}
\hline \hline
Complex & $F_{60}$ & 8.6 &  No. of  \\
...
\hline
\end{tabular}
\end{table}

The second method produces...
\end{appendix}
%
%
\end{document}

%
%-------------------------------------------------------------
%          For the appendices, table longer than a single page
%-------------------------------------------------------------

% Table will be print automatically at the end of the document, 
% after the whole appendices

\begin{appendix} %First appendix
\section{Background galaxy number counts and shear noise-levels}

% In the appendices do not forget to put the counter of the table 
% as an option

\longtab[1]{
\begin{longtable}{lrcrrrrrrrrl}
\caption{Line data and abundances ...}\\
\hline
\hline
Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & N & e &  rad & $\delta$ & $\delta$ 
red & References \\
\hline
\endfirsthead
\caption{Continued.} \\
\hline
Def & mol & Ion & $\lambda$ & $\chi$ & $\log gf$ & B & C &  rad & $\delta$ & $\delta$ 
red & References \\
\hline
\endhead
\hline
\endfoot
\hline
\endlastfoot
A & CH & 1 &3638 & 0.002 & $-$2.551 &  &  &  & $-$150 & 150 &  Jorgensen et al. (1996) \\                    
\end{longtable}
}% End longtab
\end{appendix}

%-------------------------------------------------------------
%                   For appendices and landscape, large table:
%                    in the preamble, use: \usepackage{lscape}
%-------------------------------------------------------------

\begin{appendix} %First appendix
%
\longtab[1]{
\begin{landscape}
\begin{longtable}{lrcrrrrrrrrl}
...
\end{longtable}
\end{landscape}
}% End longtab
\end{appendix}

%%%% End of aa.dem

### 2020-11-21 paper/aa-package/paper.tex
start adding biblio

@@
 \documentclass{aa}
+
+\usepackage[]{natbib}
 
@@
 \section{Introduction}
-
-   In the \emph{nucleated instability\/} (also called core
-   instability) hypothesis of giant planet
-   formation, a critical mass for static core  envelope
-   protoplanets has been found. \citet{mizuno} determined
-   the critical mass of the core to be about $12 \,M_\oplus$
-   ($M_\oplus=5.975 \times 10^{27}\,\mathrm{g}$ is the Earth mass), which
-   is independent of the outer boundary
-   conditions and therefore independent of the location in the
-   solar nebula. This critical value for the core mass corresponds
-   closely to the cores of today's giant planets.
-
-   Although no hydrodynamical study has been available many workers
-   conjectured that a collapse or rapid contraction will ensue
-   after accumulating the critical mass. The main motivation for
-   this article
-   is to investigate the stability of the static envelope at the
-   critical mass. With this aim the local, linear stability of static
-   radiative gas  spheres is investigated on the basis of Baker's
-   (\citeyear{baker}) standard one-zone model.
-
-   Phenomena similar to the ones described above for giant planet
-   formation have been found in hydrodynamical models concerning
-   star formation where protostellar cores explode
-   (Tscharnuter \citeyear{tscharnuter}, Balluch \citeyear{balluch}),
-   whereas earlier studies found quasi-steady collapse flows. The
-   similarities in the (micro)physics, i.e., constitutive relations of
-   protostellar cores and protogiant planets serve as a further
-   motivation for this study.
+The Hubble parameter $H_0$ quantifies the expansion rate of the Universe and can be measured in different ways.
+The Planck collaboration \cite{planck} has found $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$ from the Cosmic Microwave Background anisotropies spectrum \footnote{Assuming the $\Lambda$CDM cosmological model.}, whereas, cosmic distance ladder observations \cite{cepheid_H0} give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.\\ As first pointed out in \cite{Refsdal1964}, gravitational lensing offers an alternative way of determining the Hubble parameter.
+The light rays coming from a distant source, e.g.
+a quasar, can be deflected by the gravitational field of an intervening massive object, like a galaxy.
+If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, one can obtain a so-called \emph{light curve}.
+The light curves associated with different images will exhibit a mutual time-delay, due to the different paths that the photons have travelled.
+As shown in \cite{Refsdal1964}, this time-delay can be directly related to $H_0$.
+The H0LiCOW collaboration \cite{H0licow_XIII} found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL \cite{Cosmograil2020} project, enhancing the tension up to $5.3 \sigma$, and thus motivating further studies aimed at improving the precision in the estimate of $H_0$.
+\\
+Time-delays between light curves of Gravitationally Lensed Quasars (GLQs) can be estimated with various methods \cite{Tewes2013}, such as free-knot spline interpolation \cite{free} or via Gaussian Process (GP) regression \cite{Rasmussen}.
+Such methods work under the assumption that multiple images of the same quasar are fully resolved so that a light curve can be extracted for each of them.
+However, most known GLQs \cite{quasars_db} have small angular separations $\Delta\theta < \sim{3}{arcsec}$.
+This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent.
+%This would make big telescopes, with high angular resolution, the ideal instruments to perform lensed quasars monitoring.
+Furthermore, these telescopes are placed in geographical areas where effects of atmospheric turbulence are less prominent.
+However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years \cite{Cosmograil2020}.
+Therefore, due to the amount of observational requests that big telescopes have to fulfill, they can not be employed for these purposes.
+On the other hand, small/medium sized telescopes ($1$-$2$m) \cite{Borgeest1996} can be used.
+Unfortunately their reduced angular resolution, together with their often less privileged geographical positions in terms of atmospheric seeing, may worsen the effective angular resolution up to $3 arcsec$ \cite{karttunen2016fundamental}.
+In addition, small telescopes cannot observe faint distant sources.
+These shortcomings prevent the acquisition of a large number of GLQ datasets suitable for the application of the aforementioned techniques \cite{free, Rasmussen}.
+Here, we propose a novel approach based on Machine Learning (ML) algorithms to estimate the time-delay from non-resolved lensed quasar light curves.
+In particular, we train a one-dimensional Convolutional Neural Network (CNN) to map non-resolved double-lensed quasar images into the corresponding time-delay.
+Starting from open-source real data \cite{CosmograilData} of resolved quasar light curves, we combine them into a single time series to mimic the real situation in which the quasar images are not resolved.
+Then, we feed the so-obtained time series into a GP-based algorithm, which artificially generates new realistic non-resolved quasar light curves along with their corresponding time-delays.
+Preliminary experiments based on data of quasars HE0435 and HS2209 demonstrate the effectiveness of the proposed method.
 
 %--------------------------------------------------------------------
-\section{Baker's standard one-zone model}
+\section{Light Curves Simulation}
 
@@
 
-\begin{thebibliography}{}
-
-  \bibitem[Baker(1966)]{baker} Baker, N. 1966,
-      in Stellar Evolution,
-      ed.\ R. F. Stein,\& A. G. W. Cameron
-      (Plenum, New York) 333
-
-   \bibitem[Balluch(1988)]{balluch} Balluch, M. 1988,
-      A\&A, 200, 58
-
-   \bibitem[Cox(1980)]{cox} Cox, J. P. 1980,
-      Theory of Stellar Pulsation
-      (Princeton University Press, Princeton) 165
-
-   \bibitem[Cox(1969)]{cox69} Cox, A. N.,\& Stewart, J. N. 1969,
-      Academia Nauk, Scientific Information 15, 1
-
-   \bibitem[Mizuno(1980)]{mizuno} Mizuno H. 1980,
-      Prog. Theor. Phys., 64, 544
-
-   \bibitem[Tscharnuter(1987)]{tscharnuter} Tscharnuter W. M. 1987,
-      A\&A, 188, 55
-
-   \bibitem[Terlevich(1992)]{terlevich} Terlevich, R. 1992, in ASP Conf. Ser. 31, 
-      Relationships between Active Galactic Nuclei and Starburst Galaxies, 
-      ed. A. V. Filippenko, 13
-
-   \bibitem[Yorke(1980a)]{yorke80a} Yorke, H. W. 1980a,
-      A\&A, 86, 286
-
-   \bibitem[Zheng(1997)]{zheng} Zheng, W., Davidsen, A. F., Tytler, D. \& Kriss, G. A.
-      1997, preprint
-\end{thebibliography}
+\bibliographystyle{./bibtex/aa} % style aa.bst
+\bibliography{./bibtex/biblio}
 
 \end{document}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Example below of non-structurated natbib references  
-To use the v8.3 macros with this form of composition of bibliography, 
-the option "bibyear" should be added to the command line 
-"\documentclass[bibyear]{aa}".
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{thebibliography}{}
-
-  \bibitem[1966]{baker} Baker, N. 1966,
-      in Stellar Evolution,
-      ed.\ R. F. Stein,\& A. G. W. Cameron
-      (Plenum, New York) 333
-
-   \bibitem[1988]{balluch} Balluch, M. 1988,
-      A\&A, 200, 58
-
-   \bibitem[1980]{cox} Cox, J. P. 1980,
-      Theory of Stellar Pulsation
-      (Princeton University Press, Princeton) 165
-
-   \bibitem[1969]{cox69} Cox, A. N.,\& Stewart, J. N. 1969,
-      Academia Nauk, Scientific Information 15, 1
-
-   \bibitem[1980]{mizuno} Mizuno H. 1980,
-      Prog. Theor. Phys., 64, 544
-
-   \bibitem[1987]{tscharnuter} Tscharnuter W. M. 1987,
-      A\&A, 188, 55
-
-   \bibitem[1992]{terlevich} Terlevich, R. 1992, in ASP Conf. Ser. 31, 
-      Relationships between Active Galactic Nuclei and Starburst Galaxies, 
-      ed. A. V. Filippenko, 13
-
-   \bibitem[1980a]{yorke80a} Yorke, H. W. 1980a,
-      A\&A, 86, 286
-
-   \bibitem[1997]{zheng} Zheng, W., Davidsen, A. F., Tytler, D. \& Kriss, G. A.
-      1997, preprint
-\end{thebibliography}
 %

### 2020-11-23 paper/aa-package/paper.tex
improve abstract

@@
    {
-   	The analysis of Cosmic Microwave Background anisotropies and the cosmic distance ladder observations have resulted in a tension on the estimated value of the Hubble parameter $H_0$.
-This has triggered the need for new methods for its determination.
-To this extent, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ based on time delay measurements of six gravitational lensed quasars.
-Such value is compatible with the distance ladder observations, increasing the tension at the $5.3 \sigma$ level.
+   	Early universe (EU) measurements and late universe (LU) observations have resulted in a tension on the estimated value of the Hubble parameter $H_0$.
+Time-delay cosmography offers an alternative method to measure $H_0$.
+In this context, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level.
+Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the error in the $H_0$ estimation.
    	}
@@
    { 
-   	The fractional error on $H_0$, obtained with the gravitational lensing method, decreases linearly with the number of lensed systems analysed.
+   	The fractional error on $H_0$, obtained with time-delay cosmography, decreases linearly with the number of lensed systems analysed.
 Moreover, such method is highly affected by the microlensing phenomenon, which can produce changes in the actual time delay of the order of days.

### 2020-11-23 paper/aa-package/paper.tex
improve abstract

@@
    { 
-   	The fractional error on $H_0$, obtained with time-delay cosmography, decreases linearly with the number of lensed systems analysed.
-Moreover, such method is highly affected by the microlensing phenomenon, which can produce changes in the actual time delay of the order of days.
-To account for these two facts, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring of the lensed source, of the order of years.
+   	The fractional error on $H_0$ is directly related to the error in the measurement of relative time delays and it linearly decreases with the number of lensed systems analysed.
+Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring of the lensed source, of the order of years.
 This cannot be achieved with big telescopes, due to the huge amount of observational requests they have to fulfill.

### 2020-12-22 paper/aa-package/paper.tex
abstract

@@
 
-   \title{Time delay estimation in non-resolved lensed quasars}
+   \title{Time delay estimation in unresolved lensed quasars}
 
@@
 Time-delay cosmography offers an alternative method to measure $H_0$.
-In this context, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level.
+In this respect, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level.
 Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the error in the $H_0$ estimation.
@@
    { 
-   	The fractional error on $H_0$ is directly related to the error in the measurement of relative time delays and it linearly decreases with the number of lensed systems analysed.
-Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring of the lensed source, of the order of years.
+   	In time-delay cosmography, the fractional error on $H_0$ is directly related to the error of relative time delays measurements and it linearly decreases with the number of lensed systems considered.
+Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring, of the order of years.
 This cannot be achieved with big telescopes, due to the huge amount of observational requests they have to fulfill.
@@
    	In this work, we present a novel approach to estimate the time-delay in unresolved lensed quasar systems.
-Our method is further motivated by recent developments in discovering more unresolved strongly-lensed QSO systems.
+Our proposal is further motivated by recent developments in discovering more unresolved strongly-lensed QSO systems.
 	}

### 2020-12-22 paper/paper.tex
abstract

--- 2020-08-07 paper/article_2.tex
@@
 \noindent The recent tensions on the measured value of the Hubble parameter $H_0$ between cosmic microwave background radiation and astrophysical observations have triggered the need of new methods for its determination.
-In view of this, an effort has been done by the H0LiCoW collaboration to use the gravitational lensing of quasars as a probe for $H_0$ through the measurement of the time delay between the multiple images of the lensed quasar.
+In this respect, an effort has been done by the H0LiCoW collaboration to use the gravitational lensing of quasars as a probe for $H_0$ through the measurement of the time delay between the multiple images of the lensed quasar.
 Measurements of the time delay require a long term monitoring of lensed quasars, of the order of years.

### 2020-12-22 paper/aa-package/paper.tex
intro

@@
 
-\usepackage[]{natbib}
+\usepackage{natbib}
+\bibliographystyle{aa}
 
@@
 
-   \author{G. Wuchterl
+   \author{Author1
           \inst{1}
           \and
-          C. Ptolemy\inst{2}\fnmsep\thanks{Just to show the usage
+          Author2\inst{2}\fnmsep\thanks{Just to show the usage
           of the elements in the author field}
@@
 
-   \keywords{giant planet formation --
-                $\kappa$-mechanism --
-                stability of gas spheres
+   \keywords{Gravitational lensing --
+                Hubble parameter --
+                Quasars -- 
+                Galaxies --
+                Machine Learning
                }
@@
 \section{Introduction}
-The Hubble parameter $H_0$ quantifies the expansion rate of the Universe and can be measured in different ways.
-The Planck collaboration \cite{planck} has found $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$ from the Cosmic Microwave Background anisotropies spectrum \footnote{Assuming the $\Lambda$CDM cosmological model.}, whereas, cosmic distance ladder observations \cite{cepheid_H0} give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.\\ As first pointed out in \cite{Refsdal1964}, gravitational lensing offers an alternative way of determining the Hubble parameter.
-The light rays coming from a distant source, e.g.
-a quasar, can be deflected by the gravitational field of an intervening massive object, like a galaxy.
-If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, one can obtain a so-called \emph{light curve}.
-The light curves associated with different images will exhibit a mutual time-delay, due to the different paths that the photons have travelled.
-As shown in \cite{Refsdal1964}, this time-delay can be directly related to $H_0$.
-The H0LiCOW collaboration \cite{H0licow_XIII} found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL \cite{Cosmograil2020} project, enhancing the tension up to $5.3 \sigma$, and thus motivating further studies aimed at improving the precision in the estimate of $H_0$.
+The Hubble parameter $H_0$ quantifies the current expansion rate of the Universe.
+The measured value of $H_0$ from different observations have led to a tension.
+In particular, EU observations [\cite{planck}] have measured $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$, whereas, LU observations [\cite{cepheid_H0}] give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.
+\\ 
+As first pointed out in [\cite{Refsdal1964}], time-delay cosmography offers an alternative way of determining the Hubble parameter: the light rays coming from a distant source, e.g.
+a quasar, can be deflected by the gravitational field of an intervening massive object, e.g.
+a galaxy.
+If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained.
+Light curves associated with different images will exhibit a mutual time-delay ($\Delta T$), due to the different paths that the photons have travelled.
+As shown in [\cite{Refsdal1964}], this time-delay is related to the Hubble parameter as $H_0 \propto 1/\Delta T$.
+The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], which has found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$.
+However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS \cite{tdcosmo}, has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles.
+This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation.
 \\
@@
 
-\bibliographystyle{./bibtex/aa} % style aa.bst
-\bibliography{./bibtex/biblio}
+%\bibliographystyle{./aa} % style aa.bst
+%\bibliographystyle{unsrtnat}
+\bibliography{./biblio}
 
@@
 %-------------------------------------------------------------
-\end{thebibliography}
+%\end{thebibliography}
 

### 2020-12-22 paper/aa-package/paper.tex
intro

@@
 \documentclass{aa}
+
+ \usepackage{braket}
 
@@
 \\
+In this respect, the fractional error of $H_0$, for an ensemble of N Gravitationally Lensed Quasars (GLQs), is related to the uncertainties in the time-delay estimation $\sigma_{\Delta T}$, line-of-sight convergence $\sigma_{los}$ and lens surface density $\sigma_{\braket{k}}$ as
+\begin{equation}
+\frac{\sigma_H^2}{H_0^2} \sim \frac{\sigma_{\Delta T}^2 + \sigma_{los}^2 + \sigma_{\braket{k}}^2}{N}
+\end{equation}
+Therefore, increasing the sample of analysed GLQs allows to reduce the error on $H_0$.
+
 Time-delays between light curves of Gravitationally Lensed Quasars (GLQs) can be estimated with various methods \cite{Tewes2013}, such as free-knot spline interpolation \cite{free} or via Gaussian Process (GP) regression \cite{Rasmussen}.

### 2020-12-25 paper/aa-package/paper.tex
intro

--- 2020-12-22 paper/aa-package/paper.tex
@@
 \\
-Time-delays between light curves of Gravitationally Lensed Quasars (GLQs) can be estimated with various methods \cite{Tewes2013}, such as free-knot spline interpolation \cite{free} or via Gaussian Process (GP) regression \cite{Rasmussen}.
-Such methods work under the assumption that multiple images of the same quasar are fully resolved so that a light curve can be extracted for each of them.
-However, most known GLQs \cite{quasars_db} have small angular separations $\Delta\theta < \sim{3}{arcsec}$.
+In this respect, the fractional error of $H_0$, for an ensemble of N Gravitationally Lensed Quasars (GLQs), is related to the uncertainties in the time-delay estimation $\sigma_{\Delta T}$, line-of-sight convergence $\sigma_{los}$ and lens surface density $\sigma_{\braket{k}}$ as [\cite{Tie_2017}]:
+\begin{equation}
+\frac{\sigma_H^2}{H_0^2} \sim \frac{\sigma_{\Delta T}^2/\Delta T^2 + \sigma_{los}^2 + \sigma_{\braket{k}}^2}{N}
+\end{equation}
+where the first two terms are dominated by random errors and their contribution to the overall error scales as $N^{-1/2}$.
+Therefore, increasing the sample of analysed GLQs allows to reduce the error on $H_0$.
+\\
+In this paper we focus on the time delay estimation and its relative error.
+To date, a sample of about $220$ GLQs is available\footnote{https://research.ast.cam.ac.uk/lensedquasars/index.html}.
+However, only a very small subset, consisting of the brightests and with with well separated multiple images, has been used to measure $H_0$.
+The reason relies on the fact that it is easier and safer to extract information, and consequently to reduce the error on $H_0$, for such systems.
+\\
+In this respect, Fig.
+\ref{fig:qsosepmagboth} shows the distribution of the multiple images versus the maximum image separation for the known GLQs.
+Systems being part of the grey region, which represent a consistent fraction of the total sample, can be regarded as not optimal candidates, since they are less bright and/or have very small image separations.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{Figures/QSO_sep_mag_both}
+	\caption[cm]{comments}
+	\label{fig:qsosepmagboth}
+\end{figure}
+\\
+In addition, [\cite{shu2020discovering}] have presented a new method to find GLQ systems from unresolved light curves.
+Such systems have therefore very small-separation between multiple images.
 This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent.
-%This would make big telescopes, with high angular resolution, the ideal instruments to perform lensed quasars monitoring.
-Furthermore, these telescopes are placed in geographical areas where effects of atmospheric turbulence are less prominent.
-However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years \cite{Cosmograil2020}.
-Therefore, due to the amount of observational requests that big telescopes have to fulfill, they can not be employed for these purposes.
-On the other hand, small/medium sized telescopes ($1$-$2$m) \cite{Borgeest1996} can be used.
-Unfortunately their reduced angular resolution, together with their often less privileged geographical positions in terms of atmospheric seeing, may worsen the effective angular resolution up to $3 arcsec$ \cite{karttunen2016fundamental}.
+However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years [\cite{Cosmograil2020}].
+Therefore, due to the amount of observational requests that big telescopes have to fulfil, they can not be employed for these purposes.
+On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used.
+Unfortunately, their reduced angular resolution is further worsened by their often less privileged geographical positions in terms of atmospheric seeing, up to 3 arcsec \cite{karttunen2016fundamental}.
 In addition, small telescopes cannot observe faint distant sources.
-These shortcomings prevent the acquisition of a large number of GLQ datasets suitable for the application of the aforementioned techniques \cite{free, Rasmussen}.
+\\
 Here, we propose a novel approach based on Machine Learning (ML) algorithms to estimate the time-delay from non-resolved lensed quasar light curves.

### 2020-12-26 paper/aa-package/paper.tex
intro

@@
 a galaxy.
-If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained.
+If the field is strong enough, multiple images of the same source are observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained.
 Light curves associated with different images will exhibit a mutual time-delay ($\Delta T$), due to the different paths that the photons have travelled.
@@
 The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], which has found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$.
-However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS \cite{tdcosmo}, has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles.
+However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS [\cite{tdcosmo}], has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles.
 This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation.
@@
 In this paper we focus on the time delay estimation and its relative error. 
+\\
 To date, a sample of about $220$ GLQs is available\footnote{https://research.ast.cam.ac.uk/lensedquasars/index.html}.
-However, only a very small subset, consisting of the brightests and with with well separated multiple images, has been used to measure $H_0$.
-The reason relies on the fact that it is easier and safer to extract information, and consequently to reduce the error on $H_0$, for such systems.
+However, only a very small subset, with well separated multiple images, has been used to measure $H_0$.
+The reason relies on the fact that it is easier and safer to extract information from such systems, and consequently reduce the error on $H_0$.
 \\
 In this respect, Fig.
-\ref{fig:qsosepmagboth} shows the distribution of the multiple images versus the maximum image separation for the known GLQs.
-Systems being part of the grey region, which represent a consistent fraction of the total sample, can be regarded as not optimal candidates, since they are less bright and/or have very small image separations.
+\ref{fig:qsosepmagboth} (bottom) shows the magnitude of the multiple images versus the maximum image separation for the known GLQs.
+Systems being part of the grey region, which represent the 70$\%$ of the total sample (as shown in the up-right histogram), have a maximum image separation below 2 arcsec.
 \begin{figure}[h]
@@
 	\includegraphics[width=1.0\linewidth]{Figures/QSO_sep_mag_both}
-	\caption[cm]{comments}
+	\caption[cm]{Top-left: distribution of known GLQs as a function of the number of multiple images.
+Top-right: distribution of known GLQs as a function of the maximum angular separation.
+Bottom (left and right): Magnitude of the multiple images versus the maximum image separation.
+The grey region contains 70$\%$ of the total GLQ sample.}
 	\label{fig:qsosepmagboth}
@@
 \\
-In addition, [\cite{shu2020discovering}] have presented a new method to find GLQ systems from unresolved light curves.
-Such systems have therefore very small-separation between multiple images.
+In addition, [\cite{shu2020discovering}] have presented a new method to find GLQs from unresolved light curves. Such systems have therefore even smaller separation between multiple images. 
+\\
 This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent.
@@
 On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used.
-Unfortunately, their reduced angular resolution is further worsened by their often less privileged geographical positions in terms of atmospheric seeing, up to 3 arcsec \cite{karttunen2016fundamental}.
-In addition, small telescopes cannot observe faint distant sources.
+Unfortunately, their already reduced angular resolution is further worsened by their less privileged geographical positions, in terms of atmospheric seeing, which can reach 3 arcsec [\cite{karttunen2016fundamental}].
 \\
-Here, we propose a novel approach based on Machine Learning (ML) algorithms to estimate the time-delay from non-resolved lensed quasar light curves.
-In particular, we train a one-dimensional Convolutional Neural Network (CNN) to map non-resolved double-lensed quasar images into the corresponding time-delay.
+Therefore, the majority of GLQs already known, together with future discoveries, will mainly appear as a single image for small/medium-size telescopes.
+For this reason, here we propose a novel approach, based on Machine Learning (ML) algorithms, to estimate the time-delay from unresolved GLQ light curves.
+\\
+For simplicity, in this work we focus on double-lensed GLQs.
+This choice is further motivated by the fact that the majority of the already known systems are doubles, as shown in Fig.
+\ref{fig:qsosepmagboth} (top-left).
+However, [\cite{shu2020discovering}] expect to find a consistent quantity of quadruply-imaged QSOs in the future.
+Therefore, we plan to extend our work for systems with N>2 images.
+\\
+The paper is structured as follows: Sec. \ref{sec:MC} describes the Monte Carlo (MC) simulations we performed....
+
+train a one-dimensional Convolutional Neural Network (CNN) to map non-resolved double-lensed quasar images into the corresponding time-delay.
 Starting from open-source real data \cite{CosmograilData} of resolved quasar light curves, we combine them into a single time series to mimic the real situation in which the quasar images are not resolved.
@@
 %--------------------------------------------------------------------
-\section{Light Curves Simulation}
+\section{Light Curves Simulation} \label{sec:MC}
 

### 2020-12-26 paper/aa-package/paper.tex
intro

@@
 \\
-The paper is structured as follows: Sec. \ref{sec:MC} describes the Monte Carlo (MC) simulations we performed....
-
-train a one-dimensional Convolutional Neural Network (CNN) to map non-resolved double-lensed quasar images into the corresponding time-delay.
-Starting from open-source real data \cite{CosmograilData} of resolved quasar light curves, we combine them into a single time series to mimic the real situation in which the quasar images are not resolved.
-Then, we feed the so-obtained time series into a GP-based algorithm, which artificially generates new realistic non-resolved quasar light curves along with their corresponding time-delays.
-Preliminary experiments based on data of quasars HE0435 and HS2209 demonstrate the effectiveness of the proposed method.
+The paper is structured as follows: Sec. \ref{sec:MC} describes the Monte Carlo (MC) simulations we performed.... Sec 3...
 

### 2020-12-28 paper/aa-package/paper.tex
Update paper.tex

@@
 In this respect, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level.
-Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the error in the $H_0$ estimation.
+Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the uncertainty in the $H_0$ estimation.
    	}
@@
    { 
-   	In time-delay cosmography, the fractional error on $H_0$ is directly related to the error of relative time delays measurements and it linearly decreases with the number of lensed systems considered.
+   	In time-delay cosmography, the fractional uncertainty on $H_0$ is directly related to the uncertainty on relative time delays measurements and it linearly decreases with the number of lensed systems considered.
 Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring, of the order of years.
@@
 On the other hand, small/medium-sized telescopes are present in a much larger number and are often characterized by more versatile observational programs.
-However, the limited resolution capabilities of such instruments and their often not privileged geographical location prevent them from providing well-separated images of the same lensed source.
+However, the limited resolution capabilities of such instruments and their often not privileged geographical location may prevent them from providing well-separated images of the same lensed source.
    	\\
@@
 The Hubble parameter $H_0$ quantifies the current expansion rate of the Universe.
-The measured value of $H_0$ from different observations have led to a tension.
+The measured values of $H_0$ from different observations led to a tension.
 In particular, EU observations [\cite{planck}] have measured $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$, whereas, LU observations [\cite{cepheid_H0}] give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.
@@
 As shown in [\cite{Refsdal1964}], this time-delay is related to the Hubble parameter as $H_0 \propto 1/\Delta T$.
-The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], which has found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$.
-However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS [\cite{tdcosmo}], has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles.
+The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], who found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$.
+However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS [\cite{tdcosmo}], found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles.
 This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation.
@@
 However, only a very small subset, with well separated multiple images, has been used to measure $H_0$.
-The reason relies on the fact that it is easier and safer to extract information from such systems, and consequently reduce the error on $H_0$.
+The reason is that it is easier and safer to extract information from such systems, and consequently reduce the error on $H_0$.
 \\
@@
 \ref{fig:qsosepmagboth} (bottom) shows the magnitude of the multiple images versus the maximum image separation for the known GLQs.
-Systems being part of the grey region, which represent the 70$\%$ of the total sample (as shown in the up-right histogram), have a maximum image separation below 2 arcsec.
+Systems being part of the grey region, which represent 70$\%$ of the total sample (as shown in the up-right histogram), have a maximum image separation below 2 arcsec.
 \begin{figure}[h]
@@
 \\
-In addition, [\cite{shu2020discovering}] have presented a new method to find GLQs from unresolved light curves. Such systems have therefore even smaller separation between multiple images. 
+In addition, [\cite{shu2020discovering}] presented a new method to find GLQs from unresolved light curves. Such systems have therefore even smaller separation between multiple images. 
 \\
 This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent.
-However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years [\cite{Cosmograil2020}].
-Therefore, due to the amount of observational requests that big telescopes have to fulfil, they can not be employed for these purposes.
+However, because of the time scales of the intrinsic variations of the sources, such observation campaigns should last years [\cite{Cosmograil2020}].
+Therefore, due to the amount of observational requests that big telescopes have to fulfil, they hardly can be employed for these purposes.
 On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used.
-Unfortunately, their already reduced angular resolution is further worsened by their less privileged geographical positions, in terms of atmospheric seeing, which can reach 3 arcsec [\cite{karttunen2016fundamental}].
+Unfortunately, their already reduced angular resolution is further worsened by their often less privileged geographical positions, in terms of atmospheric seeing, which can reach 3 arcsec [\cite{karttunen2016fundamental}].
 \\

### 2020-12-29 paper/aa-package/paper.tex
update

@@
 	\caption[cm]{Top-left: distribution of known GLQs as a function of the number of multiple images.
-Top-right: distribution of known GLQs as a function of the maximum angular separation.
+Top-right: distribution of known GLQs as a function of the maximum image separation.
 Bottom (left and right): Magnitude of the multiple images versus the maximum image separation.
@@
 \\
-The paper is structured as follows: Sec. \ref{sec:MC} describes the Monte Carlo (MC) simulations we performed.... Sec 3...
+The paper is structured as follows: Sec.
+\ref{sec:idea_and_method description} describes the ML-based method we use for evaluating the time delay between multiple images.
+Sec.
+\ref{sec:MC} describes the Monte Carlo (MC) simulations needed to train our ML algorithm.
+Sec.
+\ref{sec:test_approach} shows the results of our method in a test dataset obtained by simulating the real systems RXJ1131.
+Finally, Sec.
+\ref{sec:real_data} shows our evaluated time delays for real systems and compare it with other estimations.
+
+\section{Time delay estimation with ML technique} \label{sec:idea_and_method description}
 
@@
 \section{Light Curves Simulation} \label{sec:MC}
+
+\section{Test approach on fake dataset} \label{sec:test_approach}
+
+\section{Results on real data} \label{sec:real_data}
 
@@
 %
-   In this section the one-zone model of \citet{baker},
-   originally used to study the Cephe{\"{\i}}d pulsation mechanism, will
-   be briefly reviewed. The resulting stability criteria will be
-   rewritten in terms of local state variables, local timescales and
-   constitutive relations.
 
-   \citet{baker} investigates the stability of thin layers in
-   self-gravitating,
-   spherical gas clouds with the following properties:
-   \begin{itemize}
-      \item hydrostatic equilibrium,
-      \item thermal equilibrium,
-      \item energy transport by grey radiation diffusion.
-   \end{itemize}
-   For the one-zone-model Baker obtains necessary conditions
-   for dynamical, secular and vibrational (or pulsational)
-   stability (Eqs.\ (34a,\,b,\,c) in Baker \citeyear{baker}). Using Baker's
-   notation:
+   params
    \[
